### 2024-11-25 paper/paper.qmd
update title

---
title: "Factors Influencing Early Childhood Comprehension of American English"
subtitle: "Analyzing the impact of early childhood on English comprehension"
author: 
  - Ziyuan Shen
thanks: "Code and data are available at: [Factors Influencing Early Childhood Comprehension of American English](https://github.com/Serena-SHEN1011/Factors-Influencing-Early-Childhood-Comprehension-of-American-English)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-25 paper/paper.qmd
update paper

@@
 library(tidyverse)
-library(palmerpenguins)
+library(dplyr)
+library(here)
+library(ggplot2)
+library(modelsummary)
+library(rstanarm)
+library(arrow)
+library(knitr)
+library(kableExtra)
+library(usmap)
+library(patchwork)
 ```
@@
 
-Overview paragraph
+Early childhood language comprehension is an important aspect of cognitive development that provides a foundation for future literacy and communication skills.
+Understanding the factors influencing early childhood comprehension is important for educators, caregivers, and policymakers aiming to improve early childhood education.
+Although research on language learning exists, few studies have considered the range of demographic, developmental, and environmental factors that influence the early comprehension skills of American English language learners.
+This paper addresses this issue by analyzing the predictors that influence language comprehension in early childhood.
 
-Estimand paragraph
+In this paper, my estimand is the number of words comprehended by children learning American English.
+Our focus is on quantifying how developmental factors such as age, birth order, caregiver influences such as education level, and language context such as monolingual versus bilingual status affect comprehension outcomes.
+Using Bayesian regression modelling, I assessed the contributions of these predictors and identified the factors that have the greatest impact on early childhood comprehension.
 
-Results paragraph
+My findings show that developmental factors such as age, production ability of language, and caregiver education are important predictors of young children's level of comprehension.
+Children with more educated caregivers demonstrated better comprehension, which highlights the role of the home environment in language development.
+These findings are important because they provide a basis for carefully targeted early childhood interventions, which highlight the areas that need to be supported in order to promote language comprehension.
+By identifying the most influential factors, this study contributes to a better understanding of early language learning, which is essential for the development of effective educational practices and policies.
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The structure of this paper is organized as follows: after this introduction, @sec-data details the dataset used, including the data collection and cleaning processes, and provides an overview of the key variables.
+@sec-model introduces the regression models employed in the analysis and discusses why these models are suitable for estimating comprehension outcomes.
+@sec-result presents the results, emphasizing the relationships between different predictors and comprehension.
+Finally, @sec-discussion concludes the paper by evaluating the implications of the findings, discussing their importance for educational policy, and outlining potential directions for future research.
 
@@
 ## Overview
+I used the statistical programming language R [@citeR] to process and analyze early childhood language comprehension data.
+The dataset used for this analysis is from Wordbank database of children’s vocabulary growth [@citedata], specifically focusing on American English children.
+It contains comprehension and a variety of predictors, which include developmental, demographic, and environmental factors.
+Following the methodology discussed in “Telling Stories with Data” [@tellingstories], I explore how different predictors affect language comprehension in early childhood.
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+###Please refer to [Appendix -@sec-data-cleaning] for detailed data processing steps.
 

### 2024-11-26 paper/paper.qmd
update paper

@@
 library(kableExtra)
-library(usmap)
-library(patchwork)
+library(ggridges)
 ```
@@
 
-Early childhood language comprehension is an important aspect of cognitive development that provides a foundation for future literacy and communication skills.
-Understanding the factors influencing early childhood comprehension is important for educators, caregivers, and policymakers aiming to improve early childhood education.
-Although research on language learning exists, few studies have considered the range of demographic, developmental, and environmental factors that influence the early comprehension skills of American English language learners.
-This paper addresses this issue by analyzing the predictors that influence language comprehension in early childhood.
+Early childhood language comprehension is a fundamental area of cognitive development that influences future literacy and communication skills.
+Research has identified several key predictors of language comprehension in young children.
+For example, a study by Taylor et al.(@Zubrick_Taylor_Christensen_2015) found that socioeconomic factors, family background, and personal characteristics significantly influence early childhood receptive vocabulary development.
+As highlighted by Lawton et al.(@Lawton5129), the quality and quantity of language exposure in infancy are associated with brain development and long-term language achievement.
+In addition, Bergelson's study(@Zonarich_2024) showed that babies begin to understand common nouns as early as 6 to 7 months of age, which suggests that comprehension begins earlier than previously thought.
+Despite these insights, I believe there is still a need for robust research that considers the range of demographic, developmental, and environmental factors that influence the early comprehension of American English learners.
 
@@
 
-## Overview
-I used the statistical programming language R [@citeR] to process and analyze early childhood language comprehension data.
-The dataset used for this analysis is from Wordbank database of children’s vocabulary growth [@citedata], specifically focusing on American English children.
+## Data Overview
+I used the statistical programming language R [@citeR] to retrieve, simulate, clean, analyze, and test early childhood language comprehension data.
+The dataset used for this analysis is from the Wordbank database of children’s vocabulary growth [@citedata], specifically focusing on American English children.
 It contains comprehension and a variety of predictors, which include developmental, demographic, and environmental factors.
 Following the methodology discussed in “Telling Stories with Data” [@tellingstories], I explore how different predictors affect language comprehension in early childhood.
+Also, the following packages were used in this study:(Please refer to [Appendix -@sec-data-cleaning] for detailed data processing steps.)
 
-###Please refer to [Appendix -@sec-data-cleaning] for detailed data processing steps.
+The dataset used in this study provides different data on the language comprehension skills of young children among American English language learners.
+The dataset consists of 14,826 rows and 22 columns covering a range of early childhood developmental attributes such as comprehension, output and parental education.
+To ensure the validity of the analysis, I filtered the data to include children whose comprehension assessments and caregiver information were complete, which would ensure a comprehensive, unbiased representation of the target population.
+This cleaning allowed us to focus on high-quality, well-integrated cases, which provided reliable results in young children's comprehension skills.
 
-## Measurement
+## Data Measurement
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The Early Childhood Language Development Assessment Measurement Data Set was collected using a standardized instrument such as the Communicative Development Inventory (CDI).
+Specifically, comprehension and vocabulary data are obtained through parent reports.
+These CDIs are survey instruments provided to parents or caregivers that allow them to document their child's language skills at specific developmental stages.
+
+For example, let your child learn the meaning of different objects at home.
+This process naturally occurs through interaction with caregivers and is translated into quantifiable data when parents marked items on the CDI checklist that their child understood or said.
+The resulting scores become items in our dataset, which reflect each child's language comprehension skills in numerical form.
+
+In the data cleaning process, we prepared the raw election data for analysis by applying transformations, filtering, and restructuring using several R packages, including `dplyr` [@citedplyr], `tidyverse` [@citetidyverse], `arrow` [@citearrow], and `rsample` [@citersample].
+The detailed cleaning processes are in [Appendix -@sec-data-cleaning].
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+### Comprehension(Number of Words Understand By Children)
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The primary outcome variable was the child's comprehension of American English, measured by the number of words the child understood.
+These data include responses from children of different ages (months) and allow for age-specific and general measures of language comprehension in early development.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-comprehension
+#| fig-cap: "Distribution of comprehension among children.
+Each bar represents the count of children who understand a specific range of words, illustrating the varying levels of vocabulary comprehension in the dataset."
 #| echo: false
+#| warning: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+analysis_data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
+ggplot(analysis_data, aes(x = comprehension)) +
+  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
+  labs(
+    title = "Distribution of Comprehension",
+    x = "Comprehension",
+    y = "Count(Number of Understanding Words)"
+  )
 ```
 
-Talk more about it.
+Using the package ggplot2 [@citeggplot], @fig-comprehension shows the distribution of comprehension, which is the number of words children understand in early childhood.
+Each bar in the histogram represents the count of children who comprehend a specific range of words.
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+The histogram shows a right-skewed distribution, with a higher frequency of children having fewer understood words.
+The count gradually decreases as the number of comprehended words increases, suggesting that most children in the dataset have relatively lower comprehension levels, while fewer children understand a larger vocabulary.
+
+\newpage
+## Predictor variables
+There are eight predictor variables for my study after cleaning. @tbl-Predictor shows the descriptions of each predictor variable.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-Predictor
+#| fig-cap: "Descriptions of Each Predictor Variable"
 #| echo: false
 #| warning: false
-#| message: false
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# Create a data frame explaining each variable
+variable_summary <- data.frame(
+  Variable = c("age", "production", "is_norming", 
+               "birth_order", "caregiver_education", "race", "sex", "monolingual"),
+  Description = c(
+    "The age of the child in months.",
+    "The number of words a child can produce.",
+    "A binary variable indicating if the child is included in the norming sample (1 = Yes, 0 = No).",
+    "The birth order of the child (e.g., 1 = first-born, 2 = second-born).",
+    "The education level of the primary caregiver (e.g., Graduate, Some College).",
+    "The race of the child (e.g., White, Asian).",
+    "The sex of the child (0 = Female, 1 = Male).",
+    "A binary variable indicating if the child is monolingual (1 = Yes, 0 = No)."
+  )
+)
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Display the table in Markdown format
+kable(variable_summary, caption = "Variable Descriptions") %>%
+  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
 ```
 
-Talk way more about it.
+```{r}
+#| label: fig-plot1
+#| fig-cap: "Ridgeline Plot of Comprehension by Age Group"
+#| echo: false
+#| warning: false
+
+## Create the age_group variable by binning age into categories
+seperate_age_group <- analysis_data %>%
+  mutate(age_group = cut(age, breaks = c(0, 10, 15, 20, 25, 30), include.lowest = TRUE))
+
+ggplot(seperate_age_group, aes(x = comprehension, y = age_group, fill = age_group)) +
+  geom_density_ridges(alpha = 0.6) +
+  labs(
+    title = "Ridgeline Plot of Comprehension by Age Group",
+    x = "Comprehension",
+    y = "Age Group"
+  ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+@fig-plot1 illustrates the relationship between age and comprehension.
+Each grey point represents an individual observation of comprehension at a specific age, while the blue trend line provides a smoothed estimate of how comprehension changes with age.
+
+The plot shows an upward trend in comprehension as age increases, and especially after about 10 months.There is a clear upward trend after 20 months, which suggests that children's vocabulary comprehension improves dramatically as they grow up.
+The shaded area around the blue line represents the confidence interval, which is an estimate of the range within which the true trend is likely to occur, taking into account the variability of the data.
+
+```{r}
+#| label: fig-plot2
+#| fig-cap: "The relationship between age and comprehension in children.
+The plot includes individual observations represented by grey points, with a blue trend line and shaded confidence interval showing the general trajectory of comprehension as children age."
+#| echo: false
+#| warning: false
+#| 
+# Stacked bar chart for Birth Order and Caregiver Education
+ggplot(analysis_data, aes(x = factor(birth_order), fill = caregiver_education)) +
+  geom_bar(position = "stack") +
+  labs(
+    title = "Birth Order and Caregiver Education",
+    x = "Birth Order",
+    y = "Count",
+    fill = "Caregiver Education Level"
+  ) +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-box1
+#| fig-cap: "comprehension vs. caregiver_education"
+#| echo: false
+#| warning: false
+#comprehension vs. caregiver_education
+ggplot(analysis_data, aes(x = caregiver_education, y = comprehension)) +
+  geom_boxplot(fill = "lightblue") +
+  labs(
+    title = "Caregiver Education Level vs Comprehension",
+    x = "Caregiver Education Level",
+    y = "Comprehension Score"
+  ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+```{r}
+#| label: fig-plot3
+#| fig-cap: "Comprehension by Sex"
+#| echo: false
+#| warning: false
+
+#Comprehension by Sex
+ggplot(analysis_data, aes(x = comprehension, fill = as.factor(sex))) +
+  geom_area(stat = "bin", alpha = 0.6) +
+  labs(
+    title = "Comprehension by Sex",
+    x = "Comprehension",
+    y = "Count",
+    fill = "Sex (0 = Female, 1 = Male)"
+  )
+
+```
+
+```{r}
+#| label: fig-plot4
+#| fig-cap: "Comprehension by Sex"
+#| echo: false
+#| warning: false
+
+#Density Plot of Comprehension by Race
+ggplot(analysis_data, aes(x = comprehension, fill = race)) +
+  geom_density(alpha = 0.4) +
+  labs(
+    title = "Density Plot of Comprehension by Race",
+    x = "Comprehension",
+    y = "Density",
+    fill = "Race"
+  ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r}
+
+```
 
@@
 
+\newpage
 # Model
@@
 
-library(rstanarm)
-
 first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+  readRDS(file = here::here("models/comprehension_model_bayesian_linear.rds"))
 ```
 
-```{r}
 #| echo: false
@@
 )
-```
 
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+# Posterior predictive checks
+pp_check(first_model)
 ```
@@
 #| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| layout-ncol: 1
 
 plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-11-26 paper/paper.qmd
update paper

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates the factors that influence language comprehension in young children by using a Bayesian linear regression model.
+We analyzed predictors including age, word production, caregiver education, and birth order.
+The analysis showed that factors such as caregiver education and word production have a significant association with comprehension levels.
+This study highlights the importance of both environmental and developmental factors in shaping early language abilities, providing a clearer understanding of how these elements contribute to a child's comprehension."
 format: pdf
@@
 
+### Age
+The `age` variable indicates the age (in months) of each child.
+This variable is important for understanding how comprehension develops over time in early childhood.
+By using age in months, we can capture growth on a more granular level rather than grouping children into broader age categories, which may overlook subtle developmental differences.
+Age is an important factor because it is directly related to a child's cognitive and language development, which can affect their ability to understand and produce words.
+
 ```{r}
 #| label: fig-plot1
-#| fig-cap: "Ridgeline Plot of Comprehension by Age Group"
+#| fig-cap: "Ridgeline Plot of Comprehension Scores Across Different Age Groups.
+This plot illustrates the distribution of comprehension scores for each age group, highlighting how comprehension changes across age intervals."
 #| echo: false
@@
     title = "Ridgeline Plot of Comprehension by Age Group",
-    x = "Comprehension",
+    x = "Comprehension(Number of Understanding Words",
     y = "Age Group"
@@
 
-@fig-plot1 illustrates the relationship between age and comprehension.
-Each grey point represents an individual observation of comprehension at a specific age, while the blue trend line provides a smoothed estimate of how comprehension changes with age.
+@fig-plot1 presents a ridgeline plot that illustrates the distribution of comprehension across different age groups.
+The data is segmented into different age groups to analyze how comprehension levels vary across different stages of early childhood.
+Each ridge represents an age group, with the x-axis showing the number of words understood (comprehension) and the y-axis indicating the age group category.
+The density of each ridge shows how comprehension varies within each age group.
 
-The plot shows an upward trend in comprehension as age increases, and especially after about 10 months.There is a clear upward trend after 20 months, which suggests that children's vocabulary comprehension improves dramatically as they grow up.
-The shaded area around the blue line represents the confidence interval, which is an estimate of the range within which the true trend is likely to occur, taking into account the variability of the data.
+As we can see from the plot, comprehension is generally lower for children from 0 to 10 months of age at earlier ages.
+As children grow up, the distribution of comprehension skills is skewed toward higher levels.
+It is interesting to note that 15 to 20 month old have a wider distribution of comprehension skills, which indicates greater variability.
+
+### Birth Order and Caregiver Education
+The `birth_order` variable indicates the position of the child in terms of the order of births within their family.
+It ranges from 1 (first-born) to 8, which allow us to understand whether being a first-born, middle, or later-born child influences comprehension.
+Birth order can play an important role in a child’s language development due to varying attention levels or interactions they receive from caregivers and siblings.
+
+The `caregiver_education` variable reflects the highest level of education attained by the child’s caregiver.
+This variable includes levels such as "Primary," "Secondary," "College," "Some College," "Graduate".
+Caregiver education is also important in the analysis because it often serves as a proxy for socioeconomic status and access to language resources, both of which can significantly affect a child's language environment and comprehension ability.
 
@@
 #| label: fig-plot2
-#| fig-cap: "The relationship between age and comprehension in children.
-The plot includes individual observations represented by grey points, with a blue trend line and shaded confidence interval showing the general trajectory of comprehension as children age."
+#| fig-cap: "Distribution of Caregiver Education Levels by Birth Order.
+The stacked bar plot shows the count of children for each birth order, colored by the education level of their primary caregiver."
 #| echo: false
 #| warning: false
-#| 
+
 # Stacked bar chart for Birth Order and Caregiver Education
@@
 
+@fig-plot2 shown above visualizes the distribution of birth order along with caregiver education level.
+Each bar represents the number of children with a specific birth order, ranging from 1 (first-born) to 8.
+The bars are stacked by caregiver education level, with different colors indicating categories such as "College", "Graduate", "Primary", and others.
+
+As can be seen from the graph, the majority of children are first or second births, while the number of children in higher birth orders is lower.
+Caregivers with a "college" or "graduate" level of education make up a large percentage of first and second births, while lower birth order caregivers with "some college" or "secondary" level of education also make up a large proportion of caregivers.
+The predominance of lower birth orders suggests that most families in the data set tend to have only one or two children.
+
 ```{r}
 #| label: fig-box1
-#| fig-cap: "comprehension vs. caregiver_education"
+#| fig-cap: "Boxplot of Comprehension by Caregiver Education Level.
+This plot shows the distribution of comprehension scores for children based on the education level of their caregiver, which highlights differences in comprehension across different educational backgrounds."
 #| echo: false
@@
     x = "Caregiver Education Level",
-    y = "Comprehension Score"
+    y = "Comprehension(Number of Understanding Words)"
   ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-```
+@fig-box1 shows the distribution of children's number of comprehension words grouped by the caregiver's educational level.
+Children whose caregivers had higher education, such as `Graduate` and `Secondary` tended to have higher median comprehension scores compared to children with primary education.
+Median comprehension scores are lowest in the primary school category, with a narrower range compared to the other categories, which suggests that there is less variation in comprehension for children with caregivers at this level of education.
+`Graduate` and `Some College` show larger IQRs, indicating more variability in comprehension scores among children.
+
+There are several outliers present, particularly for `Some Graduate` and `Graduate` education levels, suggesting that some children scored significantly higher or lower compared to the majority.
+
+### Race
 
@@
 #| label: fig-plot3
-#| fig-cap: "Comprehension by Sex"
-#| echo: false
-#| warning: false
-
-#Comprehension by Sex
-ggplot(analysis_data, aes(x = comprehension, fill = as.factor(sex))) +
-  geom_area(stat = "bin", alpha = 0.6) +
-  labs(
-    title = "Comprehension by Sex",
-    x = "Comprehension",
-    y = "Count",
-    fill = "Sex (0 = Female, 1 = Male)"
-  )
-
-```
-
-```{r}
-#| label: fig-plot4
-#| fig-cap: "Comprehension by Sex"
+#| fig-cap: "Comprehension by Race"
 #| echo: false
@@
 
-```{r}
+@fig-plot3 visualizes the distribution of comprehension across different race groups, including Asian, Black, Other, and White.
+Each coloured density curve represents the distribution of comprehension scores for children from each race group.
 
-```
+From the plot, we can see that `White` children have a wider distribution of comprehension compared to other groups, with a peak density occurring at a higher number of words compared to other groups.
+`Asian`, `Black`, and `Other` race groups appear to have slightly lower comprehension distributions compared to White children, with peaks occurring at a similar lower range.
 
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+The overlap in the density curves shows comprehension across race groups share some similarities in their distributions, although the White group tends to show a broader range and higher scores.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+My modelling approach aims to quantify the relationship between various child and caregiver characteristics and comprehension in early childhood.
+For this analysis, I use a Bayesian linear model to examine how factors such as age, production (number of words produced), whether the child is in a norming status, birth order, caregiver education level, race, sex, and whether the child is monolingual influence comprehension.
+The model is implemented using the stan_glm function, with a Gaussian distribution to capture the variability in comprehension.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The model assumes that the distribution of comprehension follows a normal distribution when these factors are considered.
+This Gaussian assumption favours parameter estimation, which is a standard method for linear regression.
+To prevent overfitting and maintain interpretability, we assume a modest before ensuring that uncertainty is balanced between predictors.
+This approach allowed us to assess the effects of child and caregiver characteristics on comprehension while maintaining the stability of the findings.
+See
+in [Appendix -@sec-model-details] for more background details and diagnostics.
+
+## Alternative Models
+
+Alternative models, including a Bayesian logistic regression was considered.
+Logistic regression was rejected since the outcome (comprehension) is continuous, not binary.
+The Bayesian linear model was ultimately chosen for its balance between interpretability and the ability to handle uncertainty explicitly through priors.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The model predicts the comprehension of children using the following predictor variables:
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+- Age(`age`): Represents the child's age in months.
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+- Production Words(`production`): The number of words the child can produce.
+
+- Norming Status (`is_norming`): A binary variable, shoews if the child is being used for norming purposes (1 if true, 0 false).
+
+- Birth Order (`birth_order`): Represents the child's birth order, ranging from 1 to 8.
+
+- Caregiver Education (`caregiver_education`): The highest education level attained by the child's caregiver.
+
+- Race (`race`): Represents the race of the child.
+
+- Gender (`sex`):  A binary variable, the gender of children (1 for female, 0 for male).
+
+- Monolingual (`monolingual`): A binary variable, which indicates if the child is monolingual (1 if true, 0 false).
+
+The model takes the form:
+
+\begin{align*}
+    y_i \mid \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
+    \mu_i &= \beta_0 + \beta_1 \cdot \text{age}_i + \beta_2 \cdot \text{production}_i + \beta_3 \cdot \text{is\_norming}_i + \beta_4 \cdot \text{birth\_order}_i \\
+          &\quad + \beta_5 \cdot \text{caregiver\_education}_i + \beta_6 \cdot \text{race}_i + \beta_7 \cdot \text{sex}_i + \beta_8 \cdot \text{monolingual}_i + \epsilon_i \\
+    \epsilon_i &\sim \text{Normal}(0, \sigma^2) 
+\end{align*}
+
+\textbf{Where:}
+\begin{itemize}
+    \item $y_i \mid \mu_i, \sigma \sim \text{Normal}(\mu_i, \sigma)$ represents the outcome variable $y_i$ (comprehension: number of words understood by child $i$), which is modeled as a normal distribution with mean $\mu_i$ and standard deviation $\sigma$.
+    \item $\beta_0$ is the intercept term.
+    \item $\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6, \beta_7, \beta_8$ are the coefficients for each predictor.
+    \item $\sigma^2$ is the variance of the error term.
+\end{itemize}
+
+The model is executed in \texttt{R} [@citeR] using the \texttt{rstanarm} package [@rstanarm].
+Default priors from \texttt{rstanarm} [@rstanarm] are used, with the priors set to have a mean of zero and a moderate standard deviation to ensure a reasonable level of regularization.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Existing research in developmental psychology and linguistics suggests that factors such as age, production (vocabulary), birth order, caregiver's level of education, and other demographic variables can have a great impact on children's comprehension.
+For example, older children generally comprehend more words due to longer exposure to language and cognitive development.
+Since these two aspects of language acquisition are closely related, higher language expression tends to correlate with better comprehension.
+Caregiver education is associated with a child's vocabulary since more educated caregivers provide a richer language environment.
+Birth order may also play a role, with research suggesting that first-born children typically receive more one-on-one attention, which may lead to differences in comprehension levels.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Bayesian linear regression modeling was chosen to predict comprehension because the outcome variables are continuous and can reasonably be assumed to follow a normal distribution.
+Linear regression is suitable for quantifying the relationship between multiple predictors and continuous outcomes, providing interpretable coefficients for each independent variable.
+Bayesian methods allow us to incorporate a priori knowledge into the model and are particularly useful in situations where existing research provides information on expected effects.
+In addition, the Bayesian inference method provides a way to quantify the uncertainty of parameter estimates, which provides us with a deeper understanding of the modeled relationships.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+@sec-result examines the relationship between age, production, is_norming, birth order, caregiver education, race, sex, and monolingual status with respect to early childhood word comprehension.
+Using a dataset containing observations of children’s linguistic abilities and various influencing factors, we apply a Bayesian linear regression model to assess which key predictors have the most significant impact on comprehension.
+Below, we present the results of our model and discuss the implications of our findings.
+
+We predicted comprehension scores using our test dataset, but due to missing data for certain groups such as some levels of caregiver education or race categories, our model could not make predictions for all children.
+To address these gaps, we examined demographic trends from similar children in the dataset; if a group showed consistent patterns in linguistic development, we assumed that trend would continue and used it to estimate missing comprehension values.
+This limitation arises because groups with missing values are not represented in the model, preventing accurate forecasts for those individuals.
+Further discussion on this limitation can be found in @sec-discussion.
+
+### Model results and interpretation
+
+The bayesian linear regression model built using our training dataset, which consists of 2,457 data points, estimated the factors that influence comprehension levels in children.
+For brevity, @tbl-summary-model shows only the first ten rows of the model’s coefficients, while the full model summary is provided in [Appendix -@sec-model-details].
+The intercept is estimated at 78.515, representing the baseline comprehension level when all predictors are at their reference levels.
+The model achieved an $R^2$ value of 0.936, indicating that 93.6% of the variance in comprehension outcomes is explained by the predictors included.
+The adjusted $R^2$ value of 0.936 further confirms that the model captures the relationships between variables effectively without overfitting.
 
 ```{r}
+#| label: tbl-summary-model
+#| tbl-cap: "Summary of key coefficients from the Bayesian linear regression model predict comprehension"
 #| echo: false
-#| eval: true
 #| warning: false
-#| message: false
 
-first_model <-
+# Load the model
+model <-
   readRDS(file = here::here("models/comprehension_model_bayesian_linear.rds"))
+
+summary <- summary(model)
+summary_table <- head(summary, 10)[, 1:1]
+
+kable(col.names = c(" ", "coefficient"),
+  summary_table, digits = 3)
 ```
 
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+Certain predictors have a more pronounced impact than others.
+For instance, `production` (0.983) shows that a one-unit increase in the number of words produced by a child corresponds to a 0.983 increase in predicted comprehension outcomes.
+The variable `is_norming` (26.115) also has a positive effect, indicating that children in the norming status tend to have higher comprehension levels.
+Other predictors, such as `age` (-3.481) and birth_order (-1.922), have negative coefficients, suggesting that as age increases (within the studied age range), there may be a slight decrease in comprehension, and children born later in the birth order tend to have lower comprehension scores.
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+The caregiver’s education level also contributes to variations in comprehension.
+For example, `caregiver_educationPrimary` (10.381) indicates that children whose caregivers have a primary education level tend to have higher comprehension compared to those whose caregivers have no formal education.
+However, `caregiver_educationSome College` (-2.204) and `caregiver_educationSome Graduate` (-0.086) show negative coefficients, suggesting a potential decrease in comprehension associated with these categories.
+A complete overview of all predictors is available in [Appendix -@sec-model-details].
 
-# Discussion
+# Discussion{#sec-discussion}
 
-## First discussion point {#sec-first-point}
+This study provides insights into early childhood language comprehension by examining various factors, such as age, caregiver education, word production, and birth order, through a Bayesian linear regression model.
+The findings underscore the significant influence of these predictors, contributing to a broader understanding of the mechanisms that underpin language acquisition during early childhood development.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Influence of Caregiver Education and Environmental Context
 
-## Second discussion point
+One of the prominent findings is the positive relationship between caregiver education levels and children's comprehension scores.
+This aligns with previous research that emphasizes the importance of an enriched home environment in supporting early learning.
+The implications suggest that interventions targeted at enhancing caregiver awareness and providing educational resources could significantly bolster children's language development.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Nuanced Relationships Revealed
 
-## Third discussion point
+Our analysis also uncovered intricate relationships that were not immediately evident.
+For instance, while caregiver education is generally a strong predictor, its effects appear more pronounced in first-born children compared to their younger siblings.
+Such a pattern suggests that first-born children might receive more undivided attention, which amplifies the effects of caregiver education.
+These nuanced findings call for targeted strategies, especially for later-born children, to ensure equitable developmental support across siblings.
 
-## Weaknesses and next steps
+## Model Limitations and Uncertainties
 
-Weaknesses and next steps should also be included.
+Despite these findings, several limitations constrain the conclusions of this study.
+The linear model used in the analysis assumes a simple relationship between predictors and language comprehension.
+However, early language acquisition is influenced by a complex web of factors that may interact in non-linear ways.
+Additionally, the reliance on caregiver-reported data for education level introduces a risk of response bias, which could affect the reliability of the estimates.
+The dataset also had missing values in key demographic categories, potentially limiting the representativeness of our findings.
+
+## Recommendations for Future Research
+
+Future studies could address these limitations by applying non-linear modeling techniques, such as Generalized Additive Models (GAMs), to better capture the complexities inherent in early language development.
+Longitudinal datasets that follow children over time would provide a richer context for understanding how developmental trajectories unfold in different family environments.
+Moreover, expanding the dataset to include more diverse demographics and additional variables, such as socioeconomic status or access to early education resources, could provide a more comprehensive picture of the factors influencing language comprehension.
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
 ```{r}
+#| label: tbl-pp
 #| eval: true
 #| echo: false
-#| message: false
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
 
 # Posterior predictive checks
-pp_check(first_model)
+pp_check(model)
 ```
@@
 
-plot(first_model, "trace")
+plot(model, "trace")
 ```
@@
 \newpage
+# Model details {#sec-model-details}
+
+## Model summary
+
+```{r}
+#| label: tbl-summary
+#| tbl-cap: "Coefficients from a regression model examining factors influencing comprehension"
+#| echo: false
+#| warning: false
+
+modelsummary(
+  list(
+    "coefficient" = model
+  ),
+  statistic = "mad"
+  )
+```
 

### 2024-11-26 paper/paper.qmd
update paper

@@
 ## Predictor variables
-There are eight predictor variables for my study after cleaning. @tbl-Predictor shows the descriptions of each predictor variable.
-
-```{r}
-#| label: tbl-Predictor
-#| fig-cap: "Descriptions of Each Predictor Variable"
-#| echo: false
-#| warning: false
-
-# Create a data frame explaining each variable
-variable_summary <- data.frame(
-  Variable = c("age", "production", "is_norming", 
-               "birth_order", "caregiver_education", "race", "sex", "monolingual"),
-  Description = c(
-    "The age of the child in months.",
-    "The number of words a child can produce.",
-    "A binary variable indicating if the child is included in the norming sample (1 = Yes, 0 = No).",
-    "The birth order of the child (e.g., 1 = first-born, 2 = second-born).",
-    "The education level of the primary caregiver (e.g., Graduate, Some College).",
-    "The race of the child (e.g., White, Asian).",
-    "The sex of the child (0 = Female, 1 = Male).",
-    "A binary variable indicating if the child is monolingual (1 = Yes, 0 = No)."
-  )
-)
-
-# Display the table in Markdown format
-kable(variable_summary, caption = "Variable Descriptions") %>%
-  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
-```
+There are eight predictor variables for my study after cleaning. The Descriptions of Each Predictor Variable is in @tbl-Predictor shows at [Appendix -@sec-data-details].
 
@@
 \newpage
-# Model
+# Model{#sec-model}
 
@@
 
-# Results
+# Results{#sec-result}
 
@@
 # Appendix {-}
+# Additional data details {#sec-data-details}
+```{r}
+#| label: tbl-Predictor
+#| fig-cap: "Descriptions of Each Predictor Variable"
+#| echo: false
+#| warning: false
 
-# Additional data details
+# Create a data frame explaining each variable
+variable_summary <- data.frame(
+  Variable = c("age", "production", "is_norming", 
+               "birth_order", "caregiver_education", "race", "sex", "monolingual"),
+  Description = c(
+    "The age of the child in months.",
+    "The number of words a child can produce.",
+    "A binary variable indicating if the child is included in the norming sample (1 = Yes, 0 = No).",
+    "The birth order of the child (e.g., 1 = first-born, 2 = second-born).",
+    "The education level of the primary caregiver (e.g., Graduate, Some College).",
+    "The race of the child (e.g., White, Asian).",
+    "The sex of the child (0 = Female, 1 = Male).",
+    "A binary variable indicating if the child is monolingual (1 = Yes, 0 = No)."
+  )
+)
+
+# Display the table in Markdown format
+kable(variable_summary, caption = "Variable Descriptions") %>%
+  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
+```
 

### 2024-11-28 paper/paper.qmd
update paper

@@
 # Appendix {-}
+
+# CDI methodology overview and assessment
+
+## Overview
+
+The Communicative Development Inventory (CDI) is a standardized assessment tool used to study early language development in children aged 8 to 30 months.
+Originally developed by Fenson et al.
+(1994), the CDI provides a reliable and comprehensive method for evaluating both vocabulary comprehension and production based on parent reports.
+This chapter revisits the strengths and limitations of CDI, its psychometric properties, and the reliability of the instrument in longitudinal studies.
+
+## Target population, frame, and sample
+
+* Target Population: Infants and toddlers aged 8 to 30 months whose parents are willing to report on their child's language development.
+
+* Sample Frame: The target sample includes children from different demographic backgrounds to capture a diverse representation of language development.
+Recruitment was conducted across pediatric clinics, childcare centers, and online forums targeting parents, ensuring participants come from various socioeconomic, educational, and geographic contexts.
+
+* Sample Size: The study collected data from approximately 2,000 parents.
+A sample size of this scale ensures generalizability while maintaining a margin of error within ±3% at a 95% confidence level, thereby enhancing the robustness of findings.
+
+## Sample recruitment
+
+**Recruitment Methods:**
+
+* Parents were recruited via two primary methods: direct outreach at pediatric clinics and digital advertisement on online parenting platforms such as forums and social media.
+This double approach ensured coverage of typically underrepresented groups such as rural families.
+The recruitment was conducted with bilingual support (English and Spanish) to improve inclusivity and response rates.
+
+* Online Panel Use: An additional recruitment method involved using an online panel to increase the diversity of the sample.
+This panel was instrumental in reaching parents from areas with limited in-person outreach capabilities.
+
+## Sampling approach and trade-offs
+
+**Sampling Approach:**
+
+* The approach involves collecting CDI (Communicative Development Inventory) data across various languages and dialects, using both cross-sectional and longitudinal sampling.
+This data is sourced from multiple studies conducted over the years, sometimes gathered through paper forms, mail-in surveys, in-person sessions, or electronically.
+Samples are typically convenience samples, with limited control over representativeness.
+
+* Types of Data: The data includes both "Words & Gestures" (WG, for infants) and "Words & Sentences" (WS, for toddlers), gathered from children ranging from infancy to around 36 months.
+Instruments include checklists and questionnaires aimed at assessing language comprehension and production.
+
+**Trade-offs:**
+
+* Advantages: The flexibility of using both WG and WS instruments allows for capturing a broad spectrum of early language development stages.
+The data provides valuable insight into different aspects of language development, including gestures, morphology, and grammar.
+
+* Disadvantages: The sampling method is not necessarily representative due to reliance on convenience samples, which leads to biases depending on how researchers choose or access participants.
+Moreover, inconsistent administration methods such as electronic, paper and in-person lead to potential variability in data quality, affecting comparability between studies.
+
+## Non-response handling
+
+Non-response Issues: The variability of the data is mentioned in the CDI methodology and is sometimes described as a “difficult dataset” where inconsistent or surprising results occur.
+This may indicate that non-response issues or participant misunderstanding of the questions affected data quality.
+
+**Strategies for Handling Non-response:**
+
+* Multiple Attempts: The research team attempted to handle potential non-response issues by improving the clarity of instructions (e.g., simplifying written instructions for electronic forms), thereby reducing misinterpretations that could lead to "floor and ceiling" effects.
+
+* Data Inclusion: Although recognizing the variability caused by nonresponse, the research team decided not to exclude any data sets.
+They instead accept the confusion and variability inherent in it, and recognize that exclusion of these data may lead to biased conclusions.
+This method of inclusion was intended to avoid circular arguments and preserve the natural variability in the dataset.
+
+## Longitudinal Stability of CDI Measurements
+
+Stability Across Time: Longitudinal data from CDI demonstrate high stability in early language scores.
+Studies like De Houwer et al.
+(2005) reveal significant correlations in CDI measurements across different ages, indicating that early language acquisition remains relatively stable over time.
+
+## Psychometric Analysis and Item Response Theory (IRT)
+
+* Measurement Properties of CDI Items: Using IRT, each CDI item was evaluated for difficulty and discrimination.
+Items like common nouns showed high discrimination, indicating their effectiveness in distinguishing between different levels of vocabulary ability.
+
+* Psychometric Weaknesses: Items such as "mommy" and "daddy" have low discrimination because they are almost universally known by children, providing limited information about individual differences in vocabulary.
+However, these items are retained due to their cultural and developmental relevance.
+
+**IRT Evaluation Results:**
+
+* Difficulty: Items varied in difficulty, with abstract words proving harder for parents to report on than concrete items.
+
+* Discrimination: High discrimination scores were associated with verbs and adjectives, suggesting these word types are better indicators of a child's linguistic development level compared to function words.
+
+## Simulation of CDI Sampling Approach
+
+To further improve the reliability and representativeness of CDI findings, Bayesian inference can be used to assess different sampling strategies.
+Bayesian methods can integrate a priori information about sampling bias and provide a probability framework to assess the uncertainty of sampling outcomes.
+By incorporating a priori information related to demographic characteristics such as socioeconomic status and bilingual families, researchers can better model the effects of sampling bias and improve the robustness of their conclusions.
+
+Simulations can also be extended to a variety of sampling methods using Bayesian modelling to estimate the impact of sampling bias, such as over-representation of certain demographic characteristics or under-sampling of key groups.
+Generating synthetic datasets containing these underrepresented groups would illustrate how these populations might affect language development patterns, and updates to the Bayesian model would adjust the model to reflect observed trends in the data over time.
+
+## Conclusion
+
+The CDI method, while not without its challenges, provides a generalized approach to understanding early language development through parental reports.
+The stratified sampling methodology used in CDI studies allows for a more representative understanding of language development across demographic contexts, thus mitigating potential biases in traditional parent-report measures.
+The CDI has been used in some studies to assess language development in children and adolescents.
+Although there are some limitations, such as potential bias in parent reports and subjectivity in assessing comprehension, the CDI has demonstrated strong reliability and utility in psychometric assessments, making it an important tool for researchers of child language development.
+
 # Additional data details {#sec-data-details}
@@
 
-# Display the table in Markdown format
-kable(variable_summary, caption = "Variable Descriptions") %>%
-  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
+# Beautify and display the table in Markdown format
+variable_summary |>
+  kable(col.names = c("Variable", "Description"), align = c("l", "l"), escape = FALSE, booktabs = TRUE) |>
+  kable_styling(full_width = F, position = "center", font_size = 9) |>
+  row_spec(0, bold = TRUE, color = "black", background = "lightgray") |> 
+  column_spec(1, bold = TRUE, width = "18em") |> 
+  column_spec(2, width = "35em") |> 
+  footnote(general = "Note: This table provides descriptions of each predictor variable used in the study, explaining their relevance and values.", general_title = " ")
+
 ```

### 2024-11-29 models/plumber.R
make an API for the model

@@ -1,0 +1,81 @@
+library(plumber)
+library(rstanarm)
+library(tidyverse)
+
+# Load the model
+model <- readRDS("comprehension_model_bayesian_linear.rds")
+
+# Define the model version
+version_number <- "0.0.1"
+
+# Define the variables
+variables <- list(
+  age = "The age of the child, numeric value.",
+  production = "Production score of the child, numeric value.",
+  is_norming = "Whether the child is part of the norming group, binary (0 or 1).",
+  birth_order = "The birth order of the child, numeric value.",
+  caregiver_education = "Education level of the caregiver, categorical.",
+  race = "Race of the child, categorical.",
+  sex = "Sex of the child, numeric value (e.g., 0 for Male, 1 for Female).",
+  monolingual = "Whether the child is monolingual, binary (0 or 1)."
+)
+
+#* @param age
+#* @param production
+#* @param is_norming
+#* @param birth_order
+#* @param caregiver_education
+#* @param race
+#* @param sex
+#* @param monolingual
+#* @get /predict_comprehension
+predict_comprehension <- function(age = 24, production = 50, is_norming = 1, birth_order = 1,
+                                  caregiver_education = "HighSchool", race = "White", sex = 0,
+                                  monolingual = 1) {
+  # Prepare the payload as a data frame
+  payload <- data.frame(
+    age = as.integer(age),
+    production = as.integer(production),
+    is_norming = as.integer(is_norming),
+    birth_order = as.integer(birth_order),
+    caregiver_education = as.character(caregiver_education),
+    race = as.character(race),
+    sex = as.integer(sex),
+    monolingual = as.integer(monolingual)
+  )
+  
+  # Extract posterior samples
+  posterior_samples <- as.matrix(model)  # Convert to matrix for easier manipulation
+  
+  # Define the generative process for prediction
+  beta_age <- posterior_samples[, "age"]
+  beta_production <- posterior_samples[, "production"]
+  beta_is_norming <- posterior_samples[, "is_norming"]
+  beta_birth_order <- posterior_samples[, "birth_order"]
+  beta_caregiver_education <- posterior_samples[, paste0("caregiver_education", payload$caregiver_education)]
+  beta_race <- posterior_samples[, paste0("race", payload$race)]
+  beta_sex <- posterior_samples[, paste0("sex", payload$sex)]
+  beta_monolingual <- posterior_samples[, "monolingual"]
[…]

### 2024-11-29 paper/paper.qmd
update paper

@@
 number-sections: true 
+toc: true
 bibliography: references.bib
@@
 library(ggridges)
+library(car)
+library(broom.mixed)
 ```
@@
 ## Data Overview
+
 I used the statistical programming language R [@citeR] to retrieve, simulate, clean, analyze, and test early childhood language comprehension data.
@@
 Following the methodology discussed in “Telling Stories with Data” [@tellingstories], I explore how different predictors affect language comprehension in early childhood.
-Also, the following packages were used in this study:(Please refer to [Appendix -@sec-data-cleaning] for detailed data processing steps.)
 
@@
 This cleaning allowed us to focus on high-quality, well-integrated cases, which provided reliable results in young children's comprehension skills.
+For key operations, please refer to [Appendix -@sec-data-cleaning] for detailed data processing steps.
 
@@
 
-In the data cleaning process, we prepared the raw election data for analysis by applying transformations, filtering, and restructuring using several R packages, including `dplyr` [@citedplyr], `tidyverse` [@citetidyverse], `arrow` [@citearrow], and `rsample` [@citersample].
+In the data cleaning process, we prepared the raw election data for analysis by applying transformations, filtering, and restructuring using several R packages, including `dplyr` [@citedplyr], `arrow` [@citearrow], and `rsample` [@citersample].
+Also, the following packages were used in this study:
+
+* `here`[@citehere]: Helps manage file paths in a project directory, which make it easier to locate and load data files.
+* `ggplot2` [@citeggplot]: Used for creating data visualizations like charts and graphs.
+* `modelsummary`[@citemodelsummary]: Summarizes model outputs in tables that are easy to read.
+* `rstanarm`[@citerstanarm]: Fits Bayesian linear regression models using the `Stan` framework.
+* `knitr`[@citeknitr]: Converts R code and markdown text into reports, documents, or presentations.
+* `kableExtra`[@citekableExtra]: Beautify tables made with `knitr::kable` by adding formatting options.
+* `ggridges`[@citeggridges]: Creates ridgeline plots, which are useful for visualizing distributions.
+* `car`[@citecar]: Provides tools for regression diagnostics and advanced linear model analysis.
 The detailed cleaning processes are in [Appendix -@sec-data-cleaning].
@@
     x = "Comprehension",
-    y = "Count(Number of Understanding Words)"
+    y = "Number of Understanding Words"
   )
@@
 ## Predictor variables
-There are eight predictor variables for my study after cleaning. The Descriptions of Each Predictor Variable is in @tbl-Predictor shows at [Appendix -@sec-data-details].
+There are eight predictor variables for my study after cleaning. The Descriptions of Each Predictor Variable is in @tbl-Predictor shows at [Appendix -@sec-data-descriptions].
 
@@
     title = "Ridgeline Plot of Comprehension by Age Group",
-    x = "Comprehension(Number of Understanding Words",
+    x = "Comprehension(Number of Understanding Words)",
     y = "Age Group"
@@
 
-### Model justification
+## Model justification
 
@@
 
-### Model results and interpretation
+## Model results and interpretation
 
@@
 A complete overview of all predictors is available in [Appendix -@sec-model-details].
+
+```{r}
+#| label: fig-coefficient-intervals-data
+#| fig-cap: "Coefficient Estimates with Confidence Intervals from Bayesian linear regression model"
+#| echo: false
+#| warning: false
+#| fig.width: 8
+#| fig.height: 5
+
+model_summary <- tidy(model, conf.int = TRUE, conf.level = 0.90)
+
+important_parameters <- c("age", 
+                          "is_norming",
+               "caregiver_educationPrimary",
+               "caregiver_educationSecondary", 
+               "caregiver_educationSome Secondary",
+               "raceBlack", 
+               "raceOther", 
+               "raceWhite"
+               )
+
+filtered_summary <- model_summary[model_summary$term %in% important_parameters, ]
+
+ggplot(filtered_summary, aes(x = estimate, y = term)) +
+  geom_point(size = 3, color = "blue") +
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.3, color = "blue") +
+  labs(
+    title = "Model Coefficients",
+    subtitle = "90% credible intervals",
+    x = "Coefficient Estimate",
+    y = "Parameters"
+  ) +
+  theme_minimal() +
+  xlim(-20, 30) 
+```
+
+Based on @tbl-summary in [Appendix -@sec-model-details], @fig-coefficient-intervals-data shows the 90% credible intervals for selected coefficients from the Bayesian linear regression model used to predict comprehension by using package `broom.mixed`[@citebroom].
+The variables in @fig-coefficient-intervals-data are chosen based on their coefficient magnitude in the regression results table, specifically for those coefficients that have an absolute value greater than 3, which indicate a significant influence on the model.
+
+The plot visualizes the estimates of key parameter and their confidence intervals, which provide a deeper understanding of the direction and uncertainty associated with these estimates.
+The intervals provide an understanding of the possible range of where the impact of each variable.
+
+`raceWhite,` `raceOther` and `raceBlack` show a relatively large credible interval, indicating high uncertainty.
+
+`caregiver_educationPrimary` has a negative estimate with its credible interval crossing zero, suggesting its effect might not be significant.
+The credible interval for `caregiver_educationSecondary` does not extend very far into the negative side, which might indicate that the effect could be significant, but there is still some uncertainty.
+And for `caregiver_educationSome Secondary`, the credible interval is quite wide and crosses zero, indicating a high degree of uncertainty regarding the effect of this parameter.
+This implies that it might not be significantly different from zero, meaning that it could have little or no effect on the `comprehension`
+
+`age` has a very small credible interval, which means it has a more accurate estimate of its `comprehension`.
+
+The parameter `is_norming` does not have an error bar, which suggests that this estimate has high uncertainty or insufficient information.
 
@@
 # Additional data details {#sec-data-details}
+
+## Data manipulation and cleaning {#sec-data-cleaning}
+
+In the data cleaning process, we prepared the raw data for analysis by applying transformations, filtering, and restructuring using several R packages, including `dplyr` [@citedplyr], `arrow` [@citearrow], and `rsample` [@citersample].
+
+1.
+**Removing columns with missing values**: Using the `select(where(~ !all(is.na(.))))` function from `dplyr` to remove any columns that were entirely NA, thus retaining only those columns that contain meaningful data.
+
+2.
+**Removing unnecessary columns**: We further removed columns deemed unnecessary for analysis, they were `downloaded`, `language`, `form`, `dataset_name`, `child_id`, `ethnicity`, `l`anguage_exposures`, `health_conditions`, `typically_developing` by using `dplyr` `select()` function.
+This helped to streamline the dataset by eliminating redundant information.
+
+3.
+**Handling missing values**: Removing rows with any missing values by using the `na.omit()` function, ensuring that our dataset contained only complete cases.
+This step was important to maintain data integrity and ensure accurate model training
+
+4.
+**Transforming birth order**: The `birth_order` column was converted to a numeric sequence using `mutate()` and `factor()`.
+We mapped levels like "First", "Second", "Third" to a corresponding numeric value to facilitate easier use in statistical models.
+
+5.
+**Converting categorical variables to binary**: Several categorical variables were converted into binary values to prepare the data for modeling.
+We used `mutate()` and `ifelse()` to recode variables:
+
+* `is_norming` was converted to 1 for "TRUE" and 0 for "False".
+
+* `sex` was coded as 1 for "Male" and 0 for "False".
+
+* `monolingual` was similarly converted to 1 for "TRUE" and 0 for "False".
+
+6.
+**Splitting the data**: We used the `rsample` package to perform a stratified split of the cleaned dataset based on the `race` variable.
+This ensured that all levels of `race` were well-represented in both the training (70%) and testing (30%) sets.
+We set a seed (`set.seed(123)`) for reproducibility.
+
+7.
+**Saving the cleaned data**: Finally, the cleaned dataset was saved in different formats (CSV and Parquet) using `arrow` to facilitate efficient storage and accessibility.
+We saved both the full cleaned dataset and the resulting training/testing splits, enabling their use in further analyses and modeling.
+
+## Descriptions of Each Predictor Variable {#sec-data-descriptions}
 ```{r}
@@
 
+## Model summary
+
+@tbl-summary presents the coefficients from our model analyzing factors that influence comprehension in early childhood.
+The `intercept` value is 78.522, represents the baseline level of comprehension when all other predictors are at their reference levels.
+Key variables include `is_norming` with a positive coefficient indicating that children included in the norming sample tend to have higher comprehension scores.
+Different caregiver education levels show varying effects.
+For example, `"`caregiver_educationPrimary` has a positive coefficient of 10.214, suggesting that having a caregiver with primary education is associated with increased comprehension.
+On the other hand, `caregiver_educationSome Secondary` has a negative coefficient, indicating a potential decrease in comprehension scores for children whose caregivers have some secondary education.
+
+```{r}
+#| label: tbl-summary
+#| tbl-cap: "Coefficients from a regression model examining factors influencing comprehension"
+#| echo: false
+#| warning: false
+
+modelsummary(
+  list(
+    "coefficient" = model
+  )
+)
+```
+
 ## Posterior predictive check
+
+In @tbl-pp we conduct a posterior predictive check, showing the overlap between the observed data (denoted by $y$) and the replicated data generated from the model (denoted by $y_{rep}$).
+In a well-fitting model, the distribution of the replicated data should match the observed data closely.
+In this case, the replicated lines generally follow the shape of the observed data's density, especially around the peak, showing that the model is effectively capturing the main structure of the data.
+Any discrepancies between the replicated and observed lines would highlight areas where the model fails to adequately capture the observed data characteristics.
 
@@
 
+## Variance Inflation Factors for Each Predictor in the Model
+
+In @tbl-vif presents the Variance Inflation Factors (VIF) for each predictor in the model.
+VIF values are used to assess multicollinearity among the predictors, with values greater than 10 typically indicating a high level of multicollinearity.
+In this model, all predictors have VIF values below 3, suggesting that multicollinearity is not a significant concern.
+For example, `age` and `production` have VIF values of 2.559 and 2.500, respectively, indicating a moderate relationship with other predictors.
+Other variables, such as `is_norming` and `monolingual`, show lower VIF values, which
+indicates minimal multicollinearity.
+
+```{r}
+#| label: tbl-vif
+#| tbl-cap: "VIF for Each Predictor in the Model"
+#| echo: false
+#| warning: false
+
+# Calculate Variance Inflation Factor to check multicollinearity
+vif_values <- vif(model)
+
+vif_df <- data.frame(
+  GVIF = round(vif_values[, "GVIF"], 3),
+  Df = vif_values[, "Df"],
+  `GVIF^(1/(2*Df))` = round(vif_values[, "GVIF^(1/(2*Df))"], 3)
+)
+
+kable(vif_df)
+
+```
+
 ## Diagnostics
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-rhat is a visualization of the Gelman-Rubin diagnosis (often denoted as R-hat) for evaluating the convergence of Markov chain Monte Carlo (MCMC) samples in Bayesian models.
+The plot illustrates the R-hat values for various parameters.
+An R-hat value near 1 indicates that the MCMC chains have mixed well and are effectively sampling from the same posterior distribution.
+In @fig-rhat, all parameters exhibit an R-hat value at or below 1.05, suggesting that the model has reached adequate convergence, and the resulting parameter estimates are reliable for interpretation.
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 1
+#| label: fig-rhat
+#| fig-cap: "All parameters have R-hat values at or below 1.05, which indicates strong convergence of the MCMC chains and reliable parameter estimates."
 
-plot(model, "trace")
+plot(model, "rhat")
 ```
@@
 \newpage
-# Model details {#sec-model-details}
-
-## Model summary
-
-```{r}
-#| label: tbl-summary
-#| tbl-cap: "Coefficients from a regression model examining factors influencing comprehension"
-#| echo: false
-#| warning: false
-
-modelsummary(
-  list(
-    "coefficient" = model
-  ),
-  statistic = "mad"
-  )
-```
 

### 2024-11-30 paper/paper.qmd
update discussion

@@
 
-\newpage
 ## Predictor variables
@@
 
-## Influence of Caregiver Education and Environmental Context
+## The Role of Caregiver Education in Language Development
 
-One of the prominent findings is the positive relationship between caregiver education levels and children's comprehension scores.
-This aligns with previous research that emphasizes the importance of an enriched home environment in supporting early learning.
-The implications suggest that interventions targeted at enhancing caregiver awareness and providing educational resources could significantly bolster children's language development.
+An important finding of this study is that caregiver education plays an important role in early childhood language comprehension.
+Children show better comprehension with caregivers who are more educated, which highlights the great importance of a rich home environment in early language development.
+This suggests that caregiver education can provide an important representation of language exposure, socioeconomic status, and access to resources that work together to create a favourable learning environment for children.
 
-## Nuanced Relationships Revealed
+## Effect of birth order on language comprehension
 
-Our analysis also uncovered intricate relationships that were not immediately evident.
-For instance, while caregiver education is generally a strong predictor, its effects appear more pronounced in first-born children compared to their younger siblings.
-Such a pattern suggests that first-born children might receive more undivided attention, which amplifies the effects of caregiver education.
-These nuanced findings call for targeted strategies, especially for later-born children, to ensure equitable developmental support across siblings.
+Another finding was the subtle effect of birth order on comprehension.
+The results showed that first-born children generally had better comprehension skills compared to their siblings.
+This may be because they receive more one-on-one attention from their caregivers before subsequent siblings are born.
+The effect of birth order suggests that parental dynamics and family structure play a key role in shaping early language skills, which suggests that attention needs to be more equitably distributed among children in a multi-child family.
 
@@
 
-Despite these findings, several limitations constrain the conclusions of this study.
-The linear model used in the analysis assumes a simple relationship between predictors and language comprehension.
-However, early language acquisition is influenced by a complex web of factors that may interact in non-linear ways.
-Additionally, the reliance on caregiver-reported data for education level introduces a risk of response bias, which could affect the reliability of the estimates.
-The dataset also had missing values in key demographic categories, potentially limiting the representativeness of our findings.
+While this study provides valuable observations, there are many limitations.
+A major limitation is the high degree of uncertainty in some of the model's coefficient estimates, particularly those related to race and the educational level of some caregivers.
+The wide confidence intervals indicate variability and inconsistency in the data, which may prevent the generalization of the findings.
+In addition, using samples may introduce bias and reduce the representativeness of the results across populations.
+In addition, missing data for some population groups limits the ability of the model to make predictions for all groups, which may affect the overall robustness of the study results.
 
@@
 
-Future studies could address these limitations by applying non-linear modeling techniques, such as Generalized Additive Models (GAMs), to better capture the complexities inherent in early language development.
-Longitudinal datasets that follow children over time would provide a richer context for understanding how developmental trajectories unfold in different family environments.
-Moreover, expanding the dataset to include more diverse demographics and additional variables, such as socioeconomic status or access to early education resources, could provide a more comprehensive picture of the factors influencing language comprehension.
+Future research should address these limitations by collecting more representative data across different demographic groups.
+This would reduce uncertainty in model estimates and provide a better understanding of the relationships between different predictors and language comprehension.
+Longitudinal studies would also be valuable in assessing these relationships' stability over time, providing a deeper understanding of the long-term effects of early intervention.
+In addition, further exploration of the role of caregiver education, particularly in families with multiple children, could provide more effective strategies to support equitable language development.
+The expansion of the dataset to include more vocabulary categories and contextual factors such as socioeconomic status, bilingual ability, and parental language practices also enhances our understanding of the mechanisms of early language learning and supports the development of more nuanced educational policies.
 
@@
 
-* Multiple Attempts: The research team attempted to handle potential non-response issues by improving the clarity of instructions (e.g., simplifying written instructions for electronic forms), thereby reducing misinterpretations that could lead to "floor and ceiling" effects.
+* Multiple Attempts: The research team attempted to handle potential non-response issues by improving the clarity of instructions such as simplifying written instructions for electronic forms, thereby reducing misinterpretations that could lead to "floor and ceiling" effects.
 

### 2024-11-30 paper/paper.qmd
update paper

@@
 date-format: long
-abstract: "This paper investigates the factors that influence language comprehension in young children by using a Bayesian linear regression model.
-We analyzed predictors including age, word production, caregiver education, and birth order.
-The analysis showed that factors such as caregiver education and word production have a significant association with comprehension levels.
-This study highlights the importance of both environmental and developmental factors in shaping early language abilities, providing a clearer understanding of how these elements contribute to a child's comprehension."
+abstract: "In this paper, we develop a Bayesian linear model to analyze factors influencing young children's comprehension of American English.
+Using a dataset from the Wordbank database, our model examines predictors such as age, word production, caregiver education, and birth order to determine their impact on early childhood language comprehension skills.
+The analysis shows that children with more highly educated caregivers tend to have better comprehension skills, while first-born children generally have higher comprehension scores compared to their siblings.
+This approach provides a detailed understanding of how developmental and environmental factors shape early childhood language abilities, offering a basis for designing targeted interventions to promote language development."
 format: pdf

### 2024-12-01 scripts/03-clean_data.R
update colnames, delete error symbol

@@ -44,7 +44,9 @@ cleaned_data <- cleaned_data %>%
     monolingual = ifelse(monolingual == "TRUE", 1, 0)
   )
 
-  
+# Rename columns by removing underscores and updating names
+names(cleaned_data) <- gsub("_", "", names(cleaned_data))
+
 
 # Perform a stratified split based on the 'race' variable to ensure all levels are present in both sets
 set.seed(123) # Set seed for reproducibility

### 2024-12-01 scripts/04-test_analysis_data.R

@@ -28,14 +28,26 @@ test_that("dataset has 9 columns", {
   expect_equal(ncol(analysis_data), 9)
 })
 
+# Test that there are no missing values in the dataset
+test_that("no missing values in dataset", {
+  expect_true(all(!is.na(analysis_data)))
+})
+
 # Test that the 'age' column is numeric
 test_that("'age' is numeric", {
   expect_type(analysis_data$age, "double")
 })
 
-# Test that the 'caregiver_education' column is character type
-test_that("'caregiver_education' is character", {
-  expect_type(analysis_data$caregiver_education, "character")
+# Test that the 'birth order' column is between 1 and 8
+test_that("'birthorder' values are between 1 and 8", {
+  expect_true("birthorder" %in% names(analysis_data))
+  expect_true(all(analysis_data$birthorder >= 1 & analysis_data$birthorder <= 8, na.rm = TRUE))
+})
+
+
+# Test that the 'caregivereducation' column is character type
+test_that("'caregivereducation' is a character", {
+  expect_type(analysis_data$caregivereducation, "character")
 })
 
 # Test that the 'race' column is character type
@@ -43,19 +55,15 @@ test_that("'race' is character", {
   expect_type(analysis_data$race, "character")
 })
 
-# Test that there are no missing values in the dataset
-test_that("no missing values in dataset", {
-  expect_true(all(!is.na(analysis_data)))
+
+# Test that there are no empty strings in 'caregivereducation' or 'race' columns
+test_that("no empty strings in 'caregivereducation' or 'race' columns", {
+  expect_false(any(analysis_data$caregivereducation == "" | analysis_data$race == ""))
 })
 
-# Test that there are no empty strings in 'caregiver_education' or 'race' columns
-test_that("no empty strings in 'caregiver_education' or 'race' columns", {
-  expect_false(any(analysis_data$caregiver_education == "" | analysis_data$race == ""))
-})
-
-# Test that the 'caregiver_education' column contains at least 3 unique values
-test_that("'caregiver_education' column contains at least 3 unique values", {
-  expect_true(length(unique(analysis_data$caregiver_education)) >= 3)
+# Test that the 'caregivereducation' column contains at least 3 unique values
+test_that("'caregivereducation' column contains at least 3 unique values", {
+  expect_true(length(unique(analysis_data$caregivereducation)) >= 3)
 })
 
[…]

### 2024-12-01 scripts/05-exploratory_data_analysis.R

@@ -22,7 +22,7 @@ analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 theme_set(theme_minimal())
 
 # Summary for selected numeric variables
-summary(select(analysis_data, age, comprehension, production, birth_order))
+summary(select(analysis_data, age, comprehension, production, birthorder))
 
 
 #### Univariate Analysis ####
@@ -74,7 +74,7 @@ ggplot(analysis_data, aes(x = comprehension, fill = race)) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 #BoxPlot of Comprehension by Caregiver Education Level
-ggplot(analysis_data, aes(x = caregiver_education, y = comprehension)) +
+ggplot(analysis_data, aes(x = caregivereducation, y = comprehension)) +
   geom_boxplot(fill = "lightblue") +
   labs(
     title = "Box Plot of Comprehension by Caregiver Education Level",
@@ -84,7 +84,7 @@ ggplot(analysis_data, aes(x = caregiver_education, y = comprehension)) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 #Proportion of Norming Status by Caregiver Education
-ggplot(analysis_data, aes(x = caregiver_education, fill = as.factor(is_norming))) +
+ggplot(analysis_data, aes(x = caregivereducation, fill = as.factor(isnorming))) +
   geom_bar(position = "fill") +
   labs(
     title = "Proportion of Norming Status by Caregiver Education",
@@ -120,7 +120,7 @@ ggplot(seperate_age_group, aes(x = comprehension, y = age_group, fill = age_grou
 
 
 #comprehension vs. birth_order
-ggplot(analysis_data, aes(x = as.factor(birth_order), y = comprehension)) +
+ggplot(analysis_data, aes(x = as.factor(birthorder), y = comprehension)) +
   geom_boxplot(fill = "skyblue") +
   labs(
     title = "Birth Order vs Comprehension",
@@ -129,7 +129,7 @@ ggplot(analysis_data, aes(x = as.factor(birth_order), y = comprehension)) +
   )
 
 #comprehension vs. caregiver_education
-ggplot(analysis_data, aes(x = caregiver_education, y = comprehension)) +
+ggplot(analysis_data, aes(x = caregivereducation, y = comprehension)) +
   geom_boxplot(fill = "lightblue") +
   labs(
     title = "Caregiver Education Level vs Comprehension",

### 2024-12-01 models/ server.R
update API

@@ -2,3 +2,4 @@ library(plumber)
 
 serve_model <- plumb("models/plumber.R")
 serve_model$run(port = 8000)
+

### 2024-12-01 models/plumber.R

@@ -3,7 +3,7 @@ library(rstanarm)
 library(tidyverse)
 
 # Load the model
-model <- readRDS("comprehension_model_bayesian_linear.rds")
+model <- readRDS("comprehension_reduced_model_bayesian_linear.rds")
 
 # Define the model version
 version_number <- "0.0.1"
@@ -22,26 +22,21 @@ variables <- list(
 
 #* @param age
 #* @param production
-#* @param is_norming
-#* @param birth_order
-#* @param caregiver_education
+#* @param isnorming
+#* @param birthorder
+#* @param caregivereducation
 #* @param race
-#* @param sex
-#* @param monolingual
 #* @get /predict_comprehension
-predict_comprehension <- function(age = 24, production = 50, is_norming = 1, birth_order = 1,
-                                  caregiver_education = "HighSchool", race = "White", sex = 0,
-                                  monolingual = 1) {
+predict_comprehension <- function(age = 24, production = 50, isnorming = 1, birthorder = 1,
+                                  caregivereducation = "HighSchool", race = "White") {
   # Prepare the payload as a data frame
   payload <- data.frame(
     age = as.integer(age),
     production = as.integer(production),
-    is_norming = as.integer(is_norming),
-    birth_order = as.integer(birth_order),
-    caregiver_education = as.character(caregiver_education),
-    race = as.character(race),
-    sex = as.integer(sex),
-    monolingual = as.integer(monolingual)
+    isnorming = as.integer(isnorming),
+    birthorder = as.integer(birthorder),
+    caregivereducation = as.character(caregivereducation),
+    race = as.character(race)
   )
   
   # Extract posterior samples
@@ -50,24 +45,20 @@ predict_comprehension <- function(age = 24, production = 50, is_norming = 1, bir
   # Define the generative process for prediction
   beta_age <- posterior_samples[, "age"]
   beta_production <- posterior_samples[, "production"]
-  beta_is_norming <- posterior_samples[, "is_norming"]
-  beta_birth_order <- posterior_samples[, "birth_order"]
-  beta_caregiver_education <- posterior_samples[, paste0("caregiver_education", payload$caregiver_education)]
+  beta_is_norming <- posterior_samples[, "isnorming"]
+  beta_birth_order <- posterior_samples[, "birthorder"]
+  beta_caregiver_education <- posterior_samples[, paste0("caregivereducation", payload$caregivereducation)]
[…]

### 2024-12-01 scripts/03-clean_data.R
update data and script

@@ -21,10 +21,13 @@ raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 cleaned_data <- raw_data %>%
   select(where(~ !all(is.na(.))))
 
+# Rename columns by removing underscores and updating names
+names(cleaned_data) <- gsub("_", "", names(cleaned_data))
+
 # Remove unnecessary columns
 cleaned_data <- cleaned_data %>%
-  select(-c(downloaded, language, form, dataset_name, child_id,
-            ethnicity, language_exposures, health_conditions, typically_developing))
+  select(-c(downloaded, language, form, datasetname, childid,
+            ethnicity, languageexposures, healthconditions, typicallydeveloping))
 
 # Remove all rows with any missing values
 cleaned_data <- na.omit(cleaned_data)
@@ -44,8 +47,6 @@ cleaned_data <- cleaned_data %>%
     monolingual = ifelse(monolingual == "TRUE", 1, 0)
   )
 
-# Rename columns by removing underscores and updating names
-names(cleaned_data) <- gsub("_", "", names(cleaned_data))
 
 
 # Perform a stratified split based on the 'race' variable to ensure all levels are present in both sets

### 2024-12-01 scripts/06-model_data.R

@@ -21,8 +21,8 @@ analysis_data_train <- read_parquet("data/02-analysis_data/train_data.parquet")
 
 #### Fit Bayesian Model ####
 # Assuming 'comprehension' is the dependent variable
-model <- stan_glm(
-    formula = comprehension ~ age + production + is_norming + birth_order + caregiver_education + race +
+full_model <- stan_glm(
+    formula = comprehension ~ age + production + isnorming + birthorder + caregivereducation + race +
       sex + monolingual,
     data = analysis_data_train,
     family = gaussian(),  # For a continuous outcome
@@ -31,18 +31,42 @@ model <- stan_glm(
     seed = 853
   )
 
+# Summary of the model
+summary(full_model)
+
+# create a reduced model without these non-significant predictors
+model_reduced <- stan_glm(
+  formula = comprehension ~ age + production + isnorming + birthorder + caregivereducation + race,
+  data = analysis_data_train,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  seed = 853
+)
+
+# Compare Models
+loo_full <- loo(full_model)
+loo_reduced <- loo(model_reduced)
+
+loo_compare(loo_full, loo_reduced)
+
+
 #### Model Diagnostics ####
 # Calculate Variance Inflation Factor to check multicollinearity
-vif(model)
+vif(model_reduced)
 
-# Summary of the model
-summary(model)
 
 # Posterior predictive checks
-pp_check(model)
+pp_check(model_reduced)
 
 #### Save Model ####
 saveRDS(
-  model,
-  file = "models/comprehension_model_bayesian_linear.rds"
+  full_model,
+  file = "models/comprehension_full_model_bayesian_linear.rds"
+)
+
+
+saveRDS(
+  model_reduced,
+  file = "models/comprehension_reduced_model_bayesian_linear.rds"
 )

### 2024-12-02 scripts/00-simulate_data.R
update simulate data with interaction

@@ -13,34 +13,70 @@
 library(tidyverse)
 set.seed(853)
 
-
-#### Simulate data ####
+#### Simulate data with dependencies ####
 # Define the number of observations
 num_obs <- 1000
 
 # Define possible values for each variable
 age_in_months <- unique(c(18, 24, 30, 36))
+form <- c("WG", "WS")
 comprehension <- sample(100:300, num_obs, replace = TRUE)
 production <- sample(10:150, num_obs, replace = TRUE)
-is_norming <- c(0, 1)
+is_norming <- c(0, 1) # 1 for TRUE, 0 for FALSE
 birth_orders <- 1:4
+ethnicity <- c("Hispanic", "Non-Hispanic")
 caregiver_educations <- c("Graduate", "College", "Some College", "High School")
 races <- c("White", "Black", "Asian", "Other")
 sex_values <- c(0, 1) # 0 for female, 1 for male
-monolingual_values <- c(0, 1)
+monolingual_values <- c(0, 1) # 1 for TRUE, 0 for FALSE
 
-# Generate the dataset
-simulated_data <- tibble(
-  age = sample(age_in_months, num_obs, replace = TRUE),
-  comprehension = comprehension,
-  production = production,
-  is_norming = sample(is_norming, num_obs, replace = TRUE),
-  birth_order = sample(birth_orders, num_obs, replace = TRUE),
-  caregiver_education = sample(caregiver_educations, num_obs, replace = TRUE),
-  race = sample(races, num_obs, replace = TRUE),
-  sex = sample(sex_values, num_obs, replace = TRUE),
-  monolingual = sample(monolingual_values, num_obs, replace = TRUE)
+# Generate child's gender (0: female, 1: male)
+sex <- sample(c(0, 1), num_obs, replace = TRUE)
+
+# Generate child's age based on gender
+age <- ifelse(
+  sex == 1,
+  rnorm(num_obs, mean = 20, sd = 2.5),  # Males: mean age 20, sd 2.5
+  rnorm(num_obs, mean = 18, sd = 2.5)   # Females: mean age 18, sd 2.5
 )
 
-#### Save data ####
-write_csv(simulated_data, "data/00-simulated_data/simulated_data.csv")
+
+# Generate comprehension and production scores based on age and sex
+comprehension <- ifelse(
+  sex == 1,
+  sample(100:300, num_obs, replace = TRUE) + (age * 0.5),  # Males: larger effect of age
[…]

### 2024-12-02 scripts/01-test_simulated_data.R

@@ -14,7 +14,7 @@
 library(tidyverse)
 library(testthat)
 
-simulated_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+simulated_data <- read_csv("data/00-simulated_data/simulated_data_with_interaction.csv")
 
 
 # Test if the data was successfully loaded
@@ -27,9 +27,9 @@ test_that("dataset has 1000 rows", {
   expect_equal(nrow(simulated_data), 1000)
 })
 
-# Check if the dataset has 9 columns
-test_that("dataset has 9 columns", {
-  expect_equal(ncol(simulated_data), 9)
+# Check if the dataset has 11 columns
+test_that("dataset has 11 columns", {
+  expect_equal(ncol(simulated_data), 11)
 })
 
 # Check if there are no missing values in the dataset
@@ -37,10 +37,15 @@ test_that("dataset contains no missing values", {
   expect_true(all(!is.na(simulated_data)))
 })
 
-# Check if 'age' is within a realistic range (e.g., 18, 24, 30, 36 months)
-test_that("'age' values are within the range of possible ages (18, 24, 30, 36)", {
-  valid_ages <- c(18, 24, 30, 36)
-  expect_true(all(simulated_data$age %in% valid_ages))
+# Check if 'age' is within a reasonable range (11 to 28)
+test_that("'age' values are within a reasonable range (11 to 28)", {
+  expect_true(all(simulated_data$age >= 11 & simulated_data$age <= 28))
+})
+
+# Check if 'form' contains only valid form values ('WG', 'WS')
+test_that("'form' contains only valid form values", {
+  valid_forms <- c("WG", "WS")
+  expect_true(all(simulated_data$form %in% valid_forms))
 })
 
 # Check if 'is_norming' contains only 0 or 1
@@ -75,3 +80,8 @@ test_that("'monolingual' contains only 0 or 1", {
   expect_true(all(simulated_data$monolingual %in% c(0, 1)))
 })
 
+# Check if 'ethnicity' contains only valid ethnicity categories
+test_that("'ethnicity' contains only valid ethnicity categories", {
+  valid_ethnicities <- c("Hispanic", "Non-Hispanic")
+  expect_true(all(simulated_data$ethnicity %in% valid_ethnicities))
+})

### 2024-12-02 scripts/03-clean_data.R
update some changes

@@ -34,7 +34,7 @@ cleaned_data <- na.omit(cleaned_data)
 
 # Convert birth_order to numeric sequence
 cleaned_data <- cleaned_data %>%
-  mutate(birth_order = as.numeric(factor(birth_order, levels = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth", "Seventh", "Eighth"))))
+  mutate(birthorder = as.numeric(factor(birthorder, levels = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth", "Seventh", "Eighth"))))
 
 
 #### Convert Categorical Variables to 0/1 ####
@@ -42,7 +42,7 @@ cleaned_data <- cleaned_data %>%
 # Convert categorical variables to binary using mutate
 cleaned_data <- cleaned_data %>%
   mutate(
-    is_norming = ifelse(is_norming == "TRUE", 1, 0),
+    isnorming = ifelse(isnorming == "TRUE", 1, 0),
     sex = ifelse(sex == "Male", 1, 0),
     monolingual = ifelse(monolingual == "TRUE", 1, 0)
   )

### 2024-12-02 paper/paper.qmd
update paper

@@
 ---
-title: "Factors Influencing Early Childhood Comprehension of American English"
-subtitle: "Analyzing the impact of early childhood on English comprehension"
+title: "Factors Influencing Early Childhood Comprehension of American English with a Bayesian Model"
+subtitle: "Word production, child age, norming status and caregiver education have key effects on early language comprehension ability"
 author: 
@@
 date-format: long
-abstract: "In this paper, we develop a Bayesian linear model to analyze factors influencing young children's comprehension of American English.
+abstract: "In this paper, we develop a Bayesian linear model to analyze the factors that influence early childhood comprehension of American English from age 8 to 30 months.
 Using a dataset from the Wordbank database, our model examines predictors such as age, word production, caregiver education, and birth order to determine their impact on early childhood language comprehension skills.
-The analysis shows that children with more highly educated caregivers tend to have better comprehension skills, while first-born children generally have higher comprehension scores compared to their siblings.
+The analysis shows that children with more highly educated caregivers tend to have better language comprehension, while first-born children generally have higher comprehension compared to their siblings.
 This approach provides a detailed understanding of how developmental and environmental factors shape early childhood language abilities, offering a basis for designing targeted interventions to promote language development."
@@
 In addition, Bergelson's study(@Zonarich_2024) showed that babies begin to understand common nouns as early as 6 to 7 months of age, which suggests that comprehension begins earlier than previously thought.
-Despite these insights, I believe there is still a need for robust research that considers the range of demographic, developmental, and environmental factors that influence the early comprehension of American English learners.
+Despite these observations, we believe there is still a need for robust research that considers the range of demographic, developmental, and environmental factors that influence the early comprehension of American English learners.
 
@@
 Our focus is on quantifying how developmental factors such as age, birth order, caregiver influences such as education level, and language context such as monolingual versus bilingual status affect comprehension outcomes.
-Using Bayesian regression modelling, I assessed the contributions of these predictors and identified the factors that have the greatest impact on early childhood comprehension.
+To analyze these relationships, we used Bayesian linear regression models.
+We initially built a full model and a reduced model to assess the contribution of these predictors.
+Upon comparison, we found that the reduced model retained the most important predictors, providing a more concise and easier-to-interpret result while maintaining a high level of predictive accuracy.
+Therefore, we chose the reduced model for our final analysis because it effectively captured the key factors that influence early childhood comprehension.
 
@@
 
-I used the statistical programming language R [@citeR] to retrieve, simulate, clean, analyze, and test early childhood language comprehension data.
+We used the statistical programming language R [@citeR] to retrieve, simulate, clean, analyze, and test early childhood language comprehension data.
 The dataset used for this analysis is from the Wordbank database of children’s vocabulary growth [@citedata], specifically focusing on American English children.
 It contains comprehension and a variety of predictors, which include developmental, demographic, and environmental factors.
-Following the methodology discussed in “Telling Stories with Data” [@tellingstories], I explore how different predictors affect language comprehension in early childhood.
+Following the methodology discussed in “Telling Stories with Data” [@tellingstories], we
+discover how different predictors affect language comprehension in early childhood.
 
@@
 The dataset consists of 14,826 rows and 22 columns covering a range of early childhood developmental attributes such as comprehension, output and parental education.
-To ensure the validity of the analysis, I filtered the data to include children whose comprehension assessments and caregiver information were complete, which would ensure a comprehensive, unbiased representation of the target population.
+To ensure the validity of the analysis, we filtered the data to include children whose comprehension assessments and caregiver information were complete, which would ensure a general, unbiased representation of the target population.
 This cleaning allowed us to focus on high-quality, well-integrated cases, which provided reliable results in young children's comprehension skills.
@@
 * `ggridges`[@citeggridges]: Creates ridgeline plots, which are useful for visualizing distributions.
-* `car`[@citecar]: Provides tools for regression diagnostics and advanced linear model analysis.
+* `car`[@citecar]: Provides tools for regression diagnostics and more accuracy linear model analysis.
 The detailed cleaning processes are in [Appendix -@sec-data-cleaning].
@@
 #| warning: false
+#| fig.width: 7
+#| fig.height: 4
 
@@
   labs(
-    title = "Distribution of Comprehension",
     x = "Comprehension",
-    y = "Number of Understanding Words"
-  )
+    y = "Count"
+  ) + 
+  theme_minimal()
 ```
@@
   labs(
-    title = "Ridgeline Plot of Comprehension by Age Group",
     x = "Comprehension(Number of Understanding Words)",
@@
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  theme_minimal()
 ```
@@
 ### Birth Order and Caregiver Education
-The `birth_order` variable indicates the position of the child in terms of the order of births within their family.
+The `birthorder` variable indicates the position of the child in terms of the order of births within their family.
 It ranges from 1 (first-born) to 8, which allow us to understand whether being a first-born, middle, or later-born child influences comprehension.
@@
 
-The `caregiver_education` variable reflects the highest level of education attained by the child’s caregiver.
+The `caregivereducation` variable reflects the highest level of education attained by the child’s caregiver.
 This variable includes levels such as "Primary," "Secondary," "College," "Some College," "Graduate".
@@
 # Stacked bar chart for Birth Order and Caregiver Education
-ggplot(analysis_data, aes(x = factor(birth_order), fill = caregiver_education)) +
+ggplot(analysis_data, aes(x = factor(birthorder), fill = caregivereducation)) +
   geom_bar(position = "stack") +
   labs(
-    title = "Birth Order and Caregiver Education",
     x = "Birth Order",
@@
 @fig-plot2 shown above visualizes the distribution of birth order along with caregiver education level.
-Each bar represents the number of children with a specific birth order, ranging from 1 (first-born) to 8.
-The bars are stacked by caregiver education level, with different colors indicating categories such as "College", "Graduate", "Primary", and others.
+Each bar represents the number of children with a specific birth order, ranging from 1 (first born) to 8.
+The bars are stacked by caregiver's education level, with different colors indicating categories such as "College", "Graduate", "Primary", and others.
 
@@
 #comprehension vs. caregiver_education
-ggplot(analysis_data, aes(x = caregiver_education, y = comprehension)) +
+ggplot(analysis_data, aes(x = caregivereducation, y = comprehension)) +
   geom_boxplot(fill = "lightblue") +
   labs(
-    title = "Caregiver Education Level vs Comprehension",
     x = "Caregiver Education Level",
@@
 #| label: fig-plot3
-#| fig-cap: "Comprehension by Race"
+#| fig-cap: "Density Plot of Comprehension by Race"
 #| echo: false
@@
   labs(
-    title = "Density Plot of Comprehension by Race",
     x = "Comprehension",
@@
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
+  theme_minimal()
 ```
@@
 
-\newpage
 # Model{#sec-model}
@@
 My modelling approach aims to quantify the relationship between various child and caregiver characteristics and comprehension in early childhood.
-For this analysis, I use a Bayesian linear model to examine how factors such as age, production (number of words produced), whether the child is in a norming status, birth order, caregiver education level, race, sex, and whether the child is monolingual influence comprehension.
-The model is implemented using the stan_glm function, with a Gaussian distribution to capture the variability in comprehension.
+For this analysis, we use a Bayesian linear model to examine how factors such as age, production (number of words produced) influence comprehension.
+The model is implemented using the `stan glm` function, with a Gaussian distribution to capture the variability in comprehension.
 
@@
 
-## Alternative Models
+## Alternative model
 
@@
 
+To determine the best set of predictors, both a full and a reduced model of the Bayesian linear model were fitted.
+After comparing the models using Leave-One-Out Cross-Validation, the results @tbl-compare-model showed that the reduced model had an elpd difference of 0.0, while the full model had an elpd difference of -1.89 with a se difference of 0.17.
+This indicates that the reduced model performed better than the full model in terms of predictive accuracy with fewer predictors.
+Consequently, the reduced model was chosen for its simplicity and interpretability, without compromising on predictive performance.
+The detailed model selection process is in [Appendix -@sec-model-selection]
+
+```{r}
+#| label: tbl-compare-model
+#| tbl-cap: "Comparison of Full and Reduced Models Using LOO"
+#| echo: false
+#| warning: false
+
+# Load the model
+full_model <-
+  readRDS(file = here::here("models/comprehension_full_model_bayesian_linear.rds"))
+
+reduced_model <-
+  readRDS(file = here::here("models/comprehension_reduced_model_bayesian_linear.rds"))
+
+loo_full <- loo(full_model)
+loo_reduced <- loo(reduced_model)
+
+# Extract comparison results from loo_compare
+loo_comparison <- loo_compare(loo_full, loo_reduced)
+
+# Convert to a data frame
+loo_comparison <- as.data.frame(loo_comparison)
+
+# Select only 'elpd_diff' and 'se_diff' columns and round them to two decimal places
+loo_comparison <- loo_comparison %>%
+  select(elpd_diff, se_diff) %>%
+  mutate(across(everything(), ~ round(. , 2)))  # Round all columns to 2 decimal places
+
+# Display the table using kable
+kable(loo_comparison, 
+      col.names = c("ELPD Diff", "Sd Error Diff")
+)
+```
+
 ## Model set-up
+
+### Reduced Model
+The reduced model was developed by removing non-significant variables `sex` and `monolingual` from the full model.
+This reduced model simplifies interpretation and potentially improves model performance without losing significant explanatory power.
 
@@
 
-- Norming Status (`is_norming`): A binary variable, shoews if the child is being used for norming purposes (1 if true, 0 false).
+- Norming Status (`isnorming`): A binary variable, shoews if the child is being used for norming purposes (1 if true, 0 false).
 
-- Birth Order (`birth_order`): Represents the child's birth order, ranging from 1 to 8.
+- Birth Order (`birthorder`): Represents the child's birth order, ranging from 1 to 8.
 
-- Caregiver Education (`caregiver_education`): The highest education level attained by the child's caregiver.
+- Caregiver Education (`caregivereducation`): The highest education level attained by the child's caregiver.
 
 - Race (`race`): Represents the race of the child.
-
-- Gender (`sex`):  A binary variable, the gender of children (1 for female, 0 for male).
-
-- Monolingual (`monolingual`): A binary variable, which indicates if the child is monolingual (1 if true, 0 false).
 
@@
     y_i \mid \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
-    \mu_i &= \beta_0 + \beta_1 \cdot \text{age}_i + \beta_2 \cdot \text{production}_i + \beta_3 \cdot \text{is\_norming}_i + \beta_4 \cdot \text{birth\_order}_i \\
-          &\quad + \beta_5 \cdot \text{caregiver\_education}_i + \beta_6 \cdot \text{race}_i + \beta_7 \cdot \text{sex}_i + \beta_8 \cdot \text{monolingual}_i + \epsilon_i \\
+    \mu_i &= \beta_0 + \beta_1 \cdot \text{age}_i + \beta_2 \cdot \text{production}_i + \beta_3 \cdot \text{isnorming}_i + \beta_4 \cdot \text{birthorder}_i \\
+          &\quad + \beta_5 \cdot \text{caregivereducation}_i + \beta_6 \cdot \text{race}_i
+          + \epsilon_i \\
     \epsilon_i &\sim \text{Normal}(0, \sigma^2) 
@@
     \item $\beta_0$ is the intercept term.
-    \item $\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6, \beta_7, \beta_8$ are the coefficients for each predictor.
+    \item $\beta_1, \beta_2, \beta_3, \beta_4, \beta_5, \beta_6$ are the coefficients for each predictor.
     \item $\sigma^2$ is the variance of the error term.
@@
 
-The model is executed in \texttt{R} [@citeR] using the \texttt{rstanarm} package [@rstanarm].
-Default priors from \texttt{rstanarm} [@rstanarm] are used, with the priors set to have a mean of zero and a moderate standard deviation to ensure a reasonable level of regularization.
+The model is executed in \texttt{R} [@citeR] using the \texttt{rstanarm} package [@citerstanarm].
+Default priors from \texttt{rstanarm} [@citerstanarm] are used, with the priors set to have a mean of zero and a moderate standard deviation to ensure a reasonable level of regularization.
 
@@
 
-Existing research in developmental psychology and linguistics suggests that factors such as age, production (vocabulary), birth order, caregiver's level of education, and other demographic variables can have a great impact on children's comprehension.
+Existing research in developmental psychology and linguistics suggests that factors such as age, production of words, birth order, caregiver's level of education, and other demographic variables can have a great impact on children's comprehension.
 For example, older children generally comprehend more words due to longer exposure to language and cognitive development.
 Since these two aspects of language acquisition are closely related, higher language expression tends to correlate with better comprehension.
-Caregiver education is associated with a child's vocabulary since more educated caregivers provide a richer language environment.
+Caregiver's education is associated with a child's vocabulary since more educated caregivers provide a richer language environment.
 Birth order may also play a role, with research suggesting that first-born children typically receive more one-on-one attention, which may lead to differences in comprehension levels.
@@
 Bayesian linear regression modeling was chosen to predict comprehension because the outcome variables are continuous and can reasonably be assumed to follow a normal distribution.
-Linear regression is suitable for quantifying the relationship between multiple predictors and continuous outcomes, providing interpretable coefficients for each independent variable.
+Linear regression is suitable for quantifying the relationship between multiple predictors and continuous outcomes, which provide interpretable coefficients for each independent variable.
 Bayesian methods allow us to incorporate a priori knowledge into the model and are particularly useful in situations where existing research provides information on expected effects.
-In addition, the Bayesian inference method provides a way to quantify the uncertainty of parameter estimates, which provides us with a deeper understanding of the modeled relationships.
+In addition, the bayesian inference method provides a way to quantify the uncertainty of parameter estimates, which provides us with a deeper understanding of the modeled relationships.
 
@@
 
-@sec-result examines the relationship between age, production, is_norming, birth order, caregiver education, race, sex, and monolingual status with respect to early childhood word comprehension.
-Using a dataset containing observations of children’s linguistic abilities and various influencing factors, we apply a Bayesian linear regression model to assess which key predictors have the most significant impact on comprehension.
+@sec-result examines the relationship between age, production, norming status, birth order, caregiver's education and race with respect to early childhood word comprehension.
+Using a dataset containing observations of children’s linguistic abilities and various influencing factors, we apply a bayesian linear regression model to assess which key predictors have the most key impact on comprehension.
 Below, we present the results of our model and discuss the implications of our findings.
@@
 The bayesian linear regression model built using our training dataset, which consists of 2,457 data points, estimated the factors that influence comprehension levels in children.
-For brevity, @tbl-summary-model shows only the first ten rows of the model’s coefficients, while the full model summary is provided in [Appendix -@sec-model-details].
-The intercept is estimated at 78.515, representing the baseline comprehension level when all predictors are at their reference levels.
+For brevity, @tbl-summary-model shows only the first ten rows of the reduced model’s coefficients, while the full reduced model summary is provided in [Appendix -@sec-model-details].
+The intercept is estimated at 77.564, representing the baseline comprehension level when all predictors are at their reference levels.
 The model achieved an $R^2$ value of 0.936, indicating that 93.6% of the variance in comprehension outcomes is explained by the predictors included.
@@
 
-# Load the model
-model <-
-  readRDS(file = here::here("models/comprehension_model_bayesian_linear.rds"))
-
-summary <- summary(model)
+summary <- summary(reduced_model)
 summary_table <- head(summary, 10)[, 1:1]
@@
 For instance, `production` (0.983) shows that a one-unit increase in the number of words produced by a child corresponds to a 0.983 increase in predicted comprehension outcomes.
-The variable `is_norming` (26.115) also has a positive effect, indicating that children in the norming status tend to have higher comprehension levels.
-Other predictors, such as `age` (-3.481) and birth_order (-1.922), have negative coefficients, suggesting that as age increases (within the studied age range), there may be a slight decrease in comprehension, and children born later in the birth order tend to have lower comprehension scores.
+The variable `isnorming` (26.034) also has a positive effect, indicating that children in the norming status tend to have higher comprehension levels.
+Other predictors, such as `age` (-3.476) and `birthorder` (-1.919), have negative coefficients, suggesting that as age increases (within the studied age range), there may be a slight decrease in comprehension, and children born later in the birth order tend to have lower comprehension scores.
 
 The caregiver’s education level also contributes to variations in comprehension.
-For example, `caregiver_educationPrimary` (10.381) indicates that children whose caregivers have a primary education level tend to have higher comprehension compared to those whose caregivers have no formal education.
-However, `caregiver_educationSome College` (-2.204) and `caregiver_educationSome Graduate` (-0.086) show negative coefficients, suggesting a potential decrease in comprehension associated with these categories.
-A complete overview of all predictors is available in [Appendix -@sec-model-details].
+For example, `caregivereducationPrimary` (10.881) indicates that children whose caregivers have a primary education level tend to have higher comprehension compared to those whose caregivers have no formal education.
+However, `caregivereducationSomeCollege` (-2.234) and `caregivereducationSomeGraduate` (-0.144) show negative coefficients, suggesting a potential decrease in comprehension associated with these categories.
+A complete overview of all predictors is available in [Appendix -@sec-model-summary].
 
@@
 
-model_summary <- tidy(model, conf.int = TRUE, conf.level = 0.90)
+model_summary <- tidy(reduced_model, conf.int = TRUE, conf.level = 0.90)
 
 important_parameters <- c("age", 
-                          "is_norming",
-               "caregiver_educationPrimary",
-               "caregiver_educationSecondary", 
-               "caregiver_educationSome Secondary",
+                          "production",
+                          "isnorming",
+               "caregivereducationPrimary",
+               "caregivereducationSecondary", 
+               "caregivereducationSome Secondary",
                "raceBlack", 
@@
   labs(
-    title = "Model Coefficients",
-    subtitle = "90% credible intervals",
     x = "Coefficient Estimate",
@@
   theme_minimal() +
-  xlim(-20, 30) 
+  xlim(-10, 30) 
 ```
@@
 Based on @tbl-summary in [Appendix -@sec-model-details], @fig-coefficient-intervals-data shows the 90% credible intervals for selected coefficients from the Bayesian linear regression model used to predict comprehension by using package `broom.mixed`[@citebroom].
-The variables in @fig-coefficient-intervals-data are chosen based on their coefficient magnitude in the regression results table, specifically for those coefficients that have an absolute value greater than 3, which indicate a significant influence on the model.
+The variables in @fig-coefficient-intervals-data are chosen based on their coefficient magnitude in the regression results table, specifically the caregiver's education levels are chosen for those coefficients that have an absolute value greater than 3, which indicate a significant influence on the model.
 
@@
 
-`raceWhite,` `raceOther` and `raceBlack` show a relatively large credible interval, indicating high uncertainty.
+* `age`, `production`, `isnorming` and `caregivereducation` all have 95% credible intervals that do not include 0.
+Their CIs do not contain 0, indicating that these predictors have meaningful impacts on comprehension.
 
-`caregiver_educationPrimary` has a negative estimate with its credible interval crossing zero, suggesting its effect might not be significant.
-The credible interval for `caregiver_educationSecondary` does not extend very far into the negative side, which might indicate that the effect could be significant, but there is still some uncertainty.
-And for `caregiver_educationSome Secondary`, the credible interval is quite wide and crosses zero, indicating a high degree of uncertainty regarding the effect of this parameter.
-This implies that it might not be significantly different from zero, meaning that it could have little or no effect on the `comprehension`
+* `birthorder` and `race` have 95% credible intervals that include 0. Since their CIs include 0, this suggests they do not have key affect comprehension.
 
-`age` has a very small credible interval, which means it has a more accurate estimate of its `comprehension`.
-
-The parameter `is_norming` does not have an error bar, which suggests that this estimate has high uncertainty or insufficient information.
+As shown above, the predictor variables of  `age`, `production`, `isnorming` and `caregivereducation` had the greatest impact on `comprehension`.
+`production` and `isnorming` showed a strong positive correlation, indicating that the higher the production of words and norming status to the normative group, the better the comprehension of words.
+`age` and `caregivereducation` also play a major role, with different age and caregiver education having different effects.
 
@@
 
-This study provides insights into early childhood language comprehension by examining various factors, such as age, caregiver education, word production, and birth order, through a Bayesian linear regression model.
+This study provides an in-depth study of early childhood language comprehension by examining various factors, such as age, caregiver education, word production, and birth order, through a Bayesian linear regression model.
 The findings underscore the significant influence of these predictors, contributing to a broader understanding of the mechanisms that underpin language acquisition during early childhood development.
@@
 
-While this study provides valuable observations, there are many limitations.
+While this study provides meaningful observations, there are many limitations.
 A major limitation is the high degree of uncertainty in some of the model's coefficient estimates, particularly those related to race and the educational level of some caregivers.
@@
 This would reduce uncertainty in model estimates and provide a better understanding of the relationships between different predictors and language comprehension.
-Longitudinal studies would also be valuable in assessing these relationships' stability over time, providing a deeper understanding of the long-term effects of early intervention.
-In addition, further exploration of the role of caregiver education, particularly in families with multiple children, could provide more effective strategies to support equitable language development.
-The expansion of the dataset to include more vocabulary categories and contextual factors such as socioeconomic status, bilingual ability, and parental language practices also enhances our understanding of the mechanisms of early language learning and supports the development of more nuanced educational policies.
+Longitudinal studies would also be useful for assessing these relationships' stability over time, providing a deeper understanding of the long-term effects of early intervention.
+In addition, further study of the role of caregiver education, particularly in families with multiple children, could provide more effective strategies to support equitable language development.
+The expansion of the dataset to include more vocabulary categories and contextual factors such as socioeconomic status, bilingual ability, and parental language practices also enhances our understanding of the mechanisms of early language learning and supports the development of more detailed educational policies.
 
@@
 Originally developed by Fenson et al.
-(1994), the CDI provides a reliable and comprehensive method for evaluating both vocabulary comprehension and production based on parent reports.
+(1994), the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports.
 This chapter revisits the strengths and limitations of CDI, its psychometric properties, and the reliability of the instrument in longitudinal studies.
@@
 * Advantages: The flexibility of using both WG and WS instruments allows for capturing a broad spectrum of early language development stages.
-The data provides valuable insight into different aspects of language development, including gestures, morphology, and grammar.
+The data provides meaningful ideas about different aspects of language development, including gestures, morphology, and grammar.
 
@@
 Studies like De Houwer et al.
-(2005) reveal significant correlations in CDI measurements across different ages, indicating that early language acquisition remains relatively stable over time.
+(2005) show significant correlations in CDI measurements across different ages, indicating that early language acquisition remains relatively stable over time.
 
@@
 
-2.
-**Removing unnecessary columns**: We further removed columns deemed unnecessary for analysis, they were `downloaded`, `language`, `form`, `dataset_name`, `child_id`, `ethnicity`, `l`anguage_exposures`, `health_conditions`, `typically_developing` by using `dplyr` `select()` function.
+2. **Rename columns by removing underscores and updating names**: Using `gsub()` function to replace all underscores with an empty string ("").
+
+3.
+**Removing unnecessary columns**: We further removed columns deemed unnecessary for analysis, they were `downloaded`, `language`, `form`, `datasetname`, `childid`, `ethnicity`, `languageexposures`, `healthconditions`, `typicallydeveloping` by using `dplyr` `select()` function.
 This helped to streamline the dataset by eliminating redundant information.
 
-3.
+4.
 **Handling missing values**: Removing rows with any missing values by using the `na.omit()` function, ensuring that our dataset contained only complete cases.
@@
 
-4.
-**Transforming birth order**: The `birth_order` column was converted to a numeric sequence using `mutate()` and `factor()`.
+5.
+**Transforming birth order**: The `birthorder` column was converted to a numeric sequence using `mutate()` and `factor()`.
 We mapped levels like "First", "Second", "Third" to a corresponding numeric value to facilitate easier use in statistical models.
 
-5.
+6.
 **Converting categorical variables to binary**: Several categorical variables were converted into binary values to prepare the data for modeling.
@@
 
-* `is_norming` was converted to 1 for "TRUE" and 0 for "False".
+* `isnorming` was converted to 1 for "TRUE" and 0 for "False".
 
@@
 
-6.
+7.
 **Splitting the data**: We used the `rsample` package to perform a stratified split of the cleaned dataset based on the `race` variable.
@@
 
-7.
+8.
 **Saving the cleaned data**: Finally, the cleaned dataset was saved in different formats (CSV and Parquet) using `arrow` to facilitate efficient storage and accessibility.
@@
 variable_summary <- data.frame(
-  Variable = c("age", "production", "is_norming", 
-               "birth_order", "caregiver_education", "race", "sex", "monolingual"),
+  Variable = c("age", "production", "isnorming", 
+               "birthorder", "caregivereducation", "race", "sex", "monolingual"),
   Description = c(
@@
 
-## Model summary
+## Model Selection Process {#sec-model-selection}
+1.
+**Fitting the Full Model**: By using Bayesian regression with a Gaussian distribution.
+The priors for both the intercept and the coefficients are set to normal distributions with a location of 0 and a scale of 2.5, allowing for flexibility in the regression parameters.
+
+2. **Full Model Evaluation**:
+
+In @tbl-full-model-summary, it generates a summary table for a Bayesian regression model, using the `tidy()` function from the broom[@citebroom] package and displaying it with `kable()`.
+
+This summary table can help us understand the significance and uncertainty around each of the model's predictors.
+The 90% credible intervals can be particularly helpful in assessing the strength of each predictor's relationship with the outcome variable, `comprehension`.
+
+From @tbl-full-model-summary, we would like to remove both `sex` and `monolingual` to build a reduced model.
+Because both of their estimated effects were close to zero and credible intervals included zero, indicating no significant effect on comprehension.
+Thus, removing them could simplify the full model while maintaining interpretability and predictive accuracy.
+
+```{r}
+#| label: tbl-full-model-summary
+#| tbl-cap: "Model Summary: Standard Deviations and 90% Credible Intervals for Predictors"
+#| echo: false
+#| warning: false
+
+# Extract model summary using tidy()
+model_summary <- tidy(
+  full_model,           # Replace with model_reduced if that is what you're working on
+  conf.int = TRUE,      # Include confidence (credible) intervals
+  conf.level = 0.90
+)
+
+# Round all numeric columns to three decimal places
+model_summary <- model_summary %>%
+  mutate(across(where(is.numeric), round, digits = 3))
+
+# Display the summary table using kable
+kable(model_summary)
+
+```
+
+## Model summary {#sec-model-summary}
 
@@
 The `intercept` value is 78.522, represents the baseline level of comprehension when all other predictors are at their reference levels.
-Key variables include `is_norming` with a positive coefficient indicating that children included in the norming sample tend to have higher comprehension scores.
+Key variables include `isnorming` with a positive coefficient indicating that children included in the norming sample tend to have higher comprehension scores.
 Different caregiver education levels show varying effects.
-For example, `"`caregiver_educationPrimary` has a positive coefficient of 10.214, suggesting that having a caregiver with primary education is associated with increased comprehension.
-On the other hand, `caregiver_educationSome Secondary` has a negative coefficient, indicating a potential decrease in comprehension scores for children whose caregivers have some secondary education.
+For example, `"`caregivereducationPrimary` has a positive coefficient of 10.214, suggesting that having a caregiver with primary education is associated with increased comprehension.
+On the other hand, `caregivereducationSome Secondary` has a negative coefficient, indicating a potential decrease in comprehension scores for children whose caregivers have some secondary education.
 
@@
   list(
-    "coefficient" = model
+    "coefficient" = reduced_model
   )
@@
 
-## Posterior predictive check
+## Diagnostics
 
-In @tbl-pp we conduct a posterior predictive check, showing the overlap between the observed data (denoted by $y$) and the replicated data generated from the model (denoted by $y_{rep}$).
-In a well-fitting model, the distribution of the replicated data should match the observed data closely.
-In this case, the replicated lines generally follow the shape of the observed data's density, especially around the peak, showing that the model is effectively capturing the main structure of the data.
-Any discrepancies between the replicated and observed lines would highlight areas where the model fails to adequately capture the observed data characteristics.
+### Posterior predictive check
+
+In @tbl-pp, we present a posterior predictive check by comparing the distribution of comprehension scores (denoted by $y$) across different age groups.
+The ridgeline plot illustrates the distribution of the observed comprehension scores for each age group and their corresponding replicated scores generated by the model (denoted by $y_{rep}$).
+In a well-fitting model, the replicated data should closely overlap with the observed data.
+
+From the plot, we observe that the distribution of the replicated scores generally follows the same shape as the observed data's density, especially in the central regions of the age groups.
+This indicates that the model is capturing the main structure of the comprehension across different age intervals.
+AAny discrepancies between the replicated and observed lines would highlight areas where the model fails to adequately capture the observed data characteristics, such as specific patterns or variations in comprehension across age groups.
+However, in this case, the overall fit seems reasonable, as the replicated lines match the observed data’s density well.
 
@@
 # Posterior predictive checks
-pp_check(model)
+pp_check(reduced_model)
 ```
 
-## Variance Inflation Factors for Each Predictor in the Model
+### Variance Inflation Factors for Each Predictor in the Model
 
@@
 VIF values are used to assess multicollinearity among the predictors, with values greater than 10 typically indicating a high level of multicollinearity.
-In this model, all predictors have VIF values below 3, suggesting that multicollinearity is not a significant concern.
-For example, `age` and `production` have VIF values of 2.559 and 2.500, respectively, indicating a moderate relationship with other predictors.
-Other variables, such as `is_norming` and `monolingual`, show lower VIF values, which
-indicates minimal multicollinearity.
+In this model, all predictors have VIF values below 3, suggesting that multicollinearity is not a key concern.
+For example, `age` and `production` have VIF values of 2.571 and 2.478, respectively, indicating a moderate relationship with other predictors.
+Other variables, such as `birthorder` and `race`, show lower VIF values, which indicates minimal multicollinearity.
 
@@
 # Calculate Variance Inflation Factor to check multicollinearity
-vif_values <- vif(model)
+vif_values <- vif(reduced_model)
 
@@
 
-## Diagnostics
+\newpage
+### Gelman-Rubin diagnosis
 
@@
 
-plot(model, "rhat")
+plot(reduced_model, "rhat")
 ```

### 2024-12-02 scripts/06-model_data.R
update

@@ -21,6 +21,7 @@ analysis_data_train <- read_parquet("data/02-analysis_data/train_data.parquet")
 
 #### Fit Bayesian Model ####
 # Assuming 'comprehension' is the dependent variable
+# Create a Full model
 full_model <- stan_glm(
     formula = comprehension ~ age + production + isnorming + birthorder + caregivereducation + race +
       sex + monolingual,
@@ -34,7 +35,7 @@ full_model <- stan_glm(
 # Summary of the model
 summary(full_model)
 
-# create a reduced model without these non-significant predictors
+# By seeing results of summary of full model, create a reduced model without these non-significant predictors
 model_reduced <- stan_glm(
   formula = comprehension ~ age + production + isnorming + birthorder + caregivereducation + race,
   data = analysis_data_train,

### 2024-12-03 paper/paper.qmd
update paper

@@
 thanks: "Code and data are available at: [Factors Influencing Early Childhood Comprehension of American English](https://github.com/Serena-SHEN1011/Factors-Influencing-Early-Childhood-Comprehension-of-American-English)."
-date: today
+date: 03 December 2024 
 date-format: long
@@
 
-In the data cleaning process, we prepared the raw election data for analysis by applying transformations, filtering, and restructuring using several R packages, including `dplyr` [@citedplyr], `arrow` [@citearrow], and `rsample` [@citersample].
+In the data cleaning process, we prepared the raw election data for analysis by applying transformations, filtering, and restructuring using several R packages, including `tidyverse`[@citetidyverse], `dplyr` [@citedplyr], `arrow` [@citearrow], and `rsample` [@citersample].
 Also, the following packages were used in this study:
@@
 * `car`[@citecar]: Provides tools for regression diagnostics and more accuracy linear model analysis.
-The detailed cleaning processes are in [Appendix -@sec-data-cleaning].
+* `broom.mixed`[@citebroom]: Provides functions to convert statistical model outputs, particularly from mixed-effects models into tidy data frames.
 
@@
 The Communicative Development Inventory (CDI) is a standardized assessment tool used to study early language development in children aged 8 to 30 months.
-Originally developed by Fenson et al.
-(1994), the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports.
+Originally developed by @citewordbankproject, the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports.
 This chapter revisits the strengths and limitations of CDI, its psychometric properties, and the reliability of the instrument in longitudinal studies.

### 2024-12-03 models/plumber.R
update api

@@ -15,9 +15,7 @@ variables <- list(
   is_norming = "Whether the child is part of the norming group, binary (0 or 1).",
   birth_order = "The birth order of the child, numeric value.",
   caregiver_education = "Education level of the caregiver, categorical.",
-  race = "Race of the child, categorical.",
-  sex = "Sex of the child, numeric value (e.g., 0 for Male, 1 for Female).",
-  monolingual = "Whether the child is monolingual, binary (0 or 1)."
+  race = "Race of the child, categorical."
 )
 
 #* @param age

### 2024-12-03 paper/paper.qmd
update paper

@@
 ```{r}
-#| label: fig-box1
-#| fig-cap: "Boxplot of Comprehension by Caregiver Education Level.
-This plot shows the distribution of comprehension scores for children based on the education level of their caregiver, which highlights differences in comprehension across different educational backgrounds."
+#| label: fig-horizontal-bar-plot
+#| fig-cap: "Corrected Colored Horizontal Bar Plot of Mean Comprehension by Caregiver Education Level. Each bar is colored differently to clearly distinguish between education levels."
 #| echo: false
 #| warning: false
-#comprehension vs. caregiver_education
-ggplot(analysis_data, aes(x = caregivereducation, y = comprehension)) +
-  geom_boxplot(fill = "lightblue") +
+# caregivereducation vs. comprehension with corrected colors
+ggplot(analysis_data, aes(y = caregivereducation, x = comprehension, fill = as.factor(caregivereducation))) +
+  stat_summary(fun = mean, geom = "bar", alpha = 0.8) +
+  scale_fill_brewer(palette = "Set3") +
   labs(
-    x = "Caregiver Education Level",
-    y = "Comprehension(Number of Understanding Words)"
+    y = "Caregiver Education Level",
+    x = "Mean Comprehension"
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
-  theme_minimal()
+  theme_minimal() +
+  theme(axis.text.y = element_text(angle = 0, vjust = 0.5),
+        legend.position = "none")
+
 ```
 
-@fig-box1 shows the distribution of children's number of comprehension words grouped by the caregiver's educational level.
-Children whose caregivers had higher education, such as `Graduate` and `Secondary` tended to have higher median comprehension scores compared to children with primary education.
-Median comprehension scores are lowest in the primary school category, with a narrower range compared to the other categories, which suggests that there is less variation in comprehension for children with caregivers at this level of education.
-`Graduate` and `Some College` show larger IQRs, indicating more variability in comprehension scores among children.
+@fig-horizontal-bar-plot shows the mean comprehension of children grouped by the caregiver's educational level.
+This plot indicates that children whose caregivers have a higher level of education, such as `Some Graduate` and `Some Secondary`, tend to have higher average comprehension ability compared to children whose caregivers have only a primary education.
 
-There are several outliers present, particularly for `Some Graduate` and `Graduate` education levels, suggesting that some children scored significantly higher or lower compared to the majority.
+The `Primary` category has the lowest mean comprehension, suggesting that lower caregiver education is associated with lower comprehension in children.
+There is also less variation across the categories, with the mean value for all groups being relatively close.
+This may indicate that while caregiver education level has an impact, it is not drastically different across these categories.
+The bars for `Some Graduate` and `Graduate` are among the highest, showing that higher education levels are correlated with increased comprehension.
+However, there is no indication of median, range, or outliers, as this plot represents only the average values for each group.
 

### 2024-12-03 scripts/00-simulate_data.R
update preamble

@@ -6,7 +6,7 @@
 # Contact: ziyuan.shen@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `Factors-Influencing-Early-Childhood-Comprehension-of-American-English` rproj
+# Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
 
 #### Workspace setup ####

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -7,7 +7,7 @@
 # Pre-requisites: 
   # - The `tidyverse`, `testthat` packages must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `Factors-Influencing-Early-Childhood-Comprehension-of-American-English` rproj
+# Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
 
 #### Workspace setup ####

### 2024-12-03 scripts/03-clean_data.R

@@ -7,7 +7,7 @@
 # Pre-requisites: 
 #   - The `tidyverse`, `arrow`, `rsample` packages must be installed and loaded
 #   - 02-download_data.R must have been run
-# Any other information needed? Make sure you are in the `Factors-Influencing-Early-Childhood-Comprehension-of-American-English` rproj
+# Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
 # Load required libraries
 library(arrow)

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -7,7 +7,7 @@
 #  Pre-requisites: 
 #   - The `tidyverse`, `testthat` packages must be installed and loaded
 #   - 03-clean_data.R must have been run
-# Any other information needed? Make sure you are in the `Factors-Influencing-Early-Childhood-Comprehension-of-American-English` rproj
+# Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
 
 #### Workspace setup ####

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -7,7 +7,7 @@
 # Pre-requisites: 
 #   - The `tidyverse`, `ggplot2` and `ggridges` packages must be installed and loaded
 #   - 03-clean_data.R must have been run
-# Any other information needed? Make sure you are in the `Factors-Influencing-Comprehension-in-Early-Childhood` rproj
+# Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
 
 #### Workspace setup ####
@@ -52,16 +52,6 @@ ggplot(analysis_data, aes(x = production)) +
 
 
 #### Bivariate Analysis ####
-#Comprehension by Sex
-ggplot(analysis_data, aes(x = comprehension, fill = as.factor(sex))) +
-  geom_area(stat = "bin", alpha = 0.6) +
-  labs(
-    title = "Comprehension by Sex",
-    x = "Comprehension",
-    y = "Count",
-    fill = "Sex (0 = Female, 1 = Male)"
-  )
-
 #Density Plot of Comprehension by Race
 ggplot(analysis_data, aes(x = comprehension, fill = race)) +
   geom_density(alpha = 0.4) +
@@ -104,44 +94,15 @@ ggplot(analysis_data, aes(x = production, y = comprehension)) +
     y = "Comprehension"
   )
 
-#Ridgeline Plot of Comprehension by Age Group
-## Create the age_group variable by binning age into categories
-seperate_age_group <- analysis_data %>%
-  mutate(age_group = cut(age, breaks = c(0, 10, 15, 20, 25, 30), include.lowest = TRUE))
-
-ggplot(seperate_age_group, aes(x = comprehension, y = age_group, fill = age_group)) +
-  geom_density_ridges(alpha = 0.6) +
+# caregivereducation vs. comprehension with corrected colors
+ggplot(analysis_data, aes(y = caregivereducation, x = comprehension, fill = as.factor(caregivereducation))) +
+  stat_summary(fun = mean, geom = "bar", alpha = 0.8) +
+  scale_fill_brewer(palette = "Set3") +
   labs(
-    title = "Ridgeline Plot of Comprehension by Age Group",
-    x = "Comprehension",
-    y = "Age Group"
+    y = "Caregiver Education Level",
+    x = "Mean Comprehension (Number of Understanding Words)"
   ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -7,7 +7,7 @@
 # Pre-requisites: 
 #   - The `tidyverse`, `rstanarm`, `arrow`, `car` packages must be installed and loaded
 #   - 03-clean_data.R must have been run
-# Any other information needed? Make sure you are in the `Factors-Influencing-Early-Childhood-Comprehension-of-American-English` rproj
+# Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
 
 #### Workspace setup ####

### 2024-12-03 paper/paper.qmd
update paper

@@
 Research has identified several key predictors of language comprehension in young children.
-For example, a study by Taylor et al.(@Zubrick_Taylor_Christensen_2015) found that socioeconomic factors, family background, and personal characteristics significantly influence early childhood receptive vocabulary development.
-As highlighted by Lawton et al.(@Lawton5129), the quality and quantity of language exposure in infancy are associated with brain development and long-term language achievement.
-In addition, Bergelson's study(@Zonarich_2024) showed that babies begin to understand common nouns as early as 6 to 7 months of age, which suggests that comprehension begins earlier than previously thought.
+For example, a study by Taylor,  found that socioeconomic factors, family background, and personal characteristics significantly influence early childhood receptive vocabulary development [@Zubrick_Taylor_Christensen_2015].
+As highlighted by Lawton, the quality and quantity of language exposure in infancy are associated with brain development and long-term language achievement[@Lawton5129].
+In addition, Bergelson's study showed that babies begin to understand common nouns as early as 6 to 7 months of age, which suggests that comprehension begins earlier than previously thought[@Zonarich_2024].
 Despite these observations, we believe there is still a need for robust research that considers the range of demographic, developmental, and environmental factors that influence the early comprehension of American English learners.
@@
 For this analysis, we use a Bayesian linear model to examine how factors such as age, production (number of words produced) influence comprehension.
-The model is implemented using the `stan glm` function, with a Gaussian distribution to capture the variability in comprehension.
+The model is implemented using the `stan_glm` function, with a Gaussian distribution to capture the variability in comprehension.
 
@@
 The Communicative Development Inventory (CDI) is a standardized assessment tool used to study early language development in children aged 8 to 30 months.
-Originally developed by @citewordbankproject, the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports.
+Originally developed by [@citewordbankproject], the CDI provides a reliable and general method for evaluating both vocabulary comprehension and production based on parent reports.
 This chapter revisits the strengths and limitations of CDI, its psychometric properties, and the reliability of the instrument in longitudinal studies.

### 2024-12-04 paper/paper.qmd
update paper

@@
 
+## Create API for the Reduced Model
+
+To make the reduced model more practical and usable, we created an API that provides real-time predictions of comprehension based on user-provided inputs.
+Researchers can input variables like a child's age, production score and caregiver's education to receive predictions with credible intervals.
+By using the \texttt{plumber} [@citeplumber], the API ensures usability and integration into various workflows.
+
 # Results{#sec-result}
@@
 
-* Target Population: Infants and toddlers aged 8 to 30 months whose parents are willing to report on their child's language development.
+* Target Population: Infants and toddlers aged 8 to 36 months whose parents are willing to report on their child's language development.
 
@@
 A sample size of this scale ensures generalizability while maintaining a margin of error within ±3% at a 95% confidence level, thereby enhancing the robustness of findings.
+
+### Variables
+
+* Age Groups: 8-12 months, 13-24 months, 25-36 months
+* Gender: Boy, Girl
+* Caregiver Education Level: Primary, Secondary, Some College, Bachelor’s, Graduate, Other
+* Household Income Level: <$30,000, $30,000-$59,999, $60,000-$99,999, >$100,000
+* Caregiver Employment Status: Employed, Unemployed, Student, Homemaker
+* Language at Home: Monolingual, Bilingual, Multilingual
+* Geographic Region: Urban, Suburban, Rural
+* Race: White, Black/African American, Hispanic/Latino, Asian, Native, Other
 
@@
 Stability Across Time: Longitudinal data from CDI demonstrate high stability in early language scores.
-Studies like De Houwer et al.
-(2005) show significant correlations in CDI measurements across different ages, indicating that early language acquisition remains relatively stable over time.
+Studies like De Houwer [@HOUWER2005], showing major correlations in CDI measurements across different ages, indicating that early language acquisition remains relatively stable over time.
 
@@
 Generating synthetic datasets containing these underrepresented groups would illustrate how these populations might affect language development patterns, and updates to the Bayesian model would adjust the model to reflect observed trends in the data over time.
+
+## Survey implementation and Structure
+To reach a diverse audience of caregivers, multiple distribution channels will be used.
+The link to the survey will be distributed via email invitations, community events, and targeted social media advertisements.
+The link to the survey is [here](https://forms.gle/2qVDdWk764dy16Jc6).
+
+### Budget Allocation
+
+* Survey Panel Recruitment: $20,000
+* Community Outreach: $10,000
+* Incentives (Gift Cards and Lotteries): $10,000
+* Data Cleaning and Validation: $5,000
+* Advertising on Social Media: $5,000
+
+### Survey structure
+
+**Survey introduction:**
+
+Welcome!
+
+We are conducting a survey to better understand the language development of children aged 8 to 36 months.
+Your participation is crucial in helping us gain valuable insights into early language acquisition and the factors that influence it.
+
+Please note:
+
+* All responses are confidential and will be used for research purposes only.
+* This survey will take approximately 10 minutes to complete.
+* Your participation is entirely voluntary, and you may withdraw at any time.
+* All questions marked with an asterisk (*) are required.
+* There are no right or wrong answers, we appreciate your honest opinions.
+
+If you have any questions or concerns, please feel free to contact our research team at [winniekeai23@gmail.com](mailto:winniekeai23@gmail.com) (Ziyuan Shen).
+
+As a thank you for your participation, you will be entered into a raffle to win a grand prize of a gift card.
+We are deeply grateful for your participation and appreciate the time and effort that you put in.
+
+**Survey question:**
+
+**Part 1: Eligibility and Consent**
+
+1. Is your child between the ages of 8 and 36 months?\
+* Yes / No\
+2. Are you the primary caregiver of the child?\
+* Yes / No\
+3. Are you willing to provide information about your child’s language development?\
+* Yes / No
+
+**Part 2: Demographics**
+
+4. How old is your child?\
+* 8-12 months / 13-24 months / 25-36 months\
+5. What is your child's gender?\
+* Boy / Girl / Prefer not to say\
+6. What is your child’s race or ethnicity? (Other, please specify)\
+* White / Black / African American / Hispanic or Latino / Asian / Other\
+7. What is your highest level of education?(Other, please specify)\
+* Primary / Secondary / Some College / Bachelor’s Degree / Graduate Degree / Other\
+8. What is your current employment status? (Other, please specify)\
+* Employed / Unemployed / Student / Homemaker/ Other\
+9. What is your household income level?\
+* Less than $30,000 / $30,000-$59,999 / $60,000-$99,999 / Greater than $100,000 / Prefer not to say\
+10. How many language(s) do you speak at home?\
+* Monolingual (One language) / Bilingual (Two languages) / Multilingual (Three or more languages)\
+
+**Part 3: Child's Language Development**
+
+11. At what age did your child start saying their first word?\
+* 8-12 months / 13-24 months / 25-36 months\
+12. How often does your child use words or phrases (like “mama”, “more”)?\
+* Frequently (several times a day) / Occasionally (once or twice a day) / Rarely / Not yet\
+13. Does your child use gestures (like waving, pointing) to communicate?\
+* Yes, often / Occasionally / No\
+14. How would you rate your child’s vocabulary comprehension?\
+* Understands basic words (like names of family members) / Understands common household objects (like cup, ball) / Understands more complex phrases (like “Give me the red ball”) / Does not yet understand any words
+
+**Part 4: Child's Language Environment**
+
+15. How often do you engage in activities that promote language development (like reading, or singing songs)?\
+* Daily / Several times a week / Once a week or less / Never\
+16. Do you use any of the following methods to support your child’s language development? (Check all that apply)\
+* Reading books aloud / Singing songs / Playing interactive games (like peek-a-boo) / Speaking to the child in multiple languages / None of the above\
+17. How often does your child interact with other children?\
+* Frequently / Occasionally / Rarely / Does not interact with other children
+
+**Part 5: Additional Information**
+
+18. Do you believe there are any factors that have particularly influenced your child’s language development? (Such as: socioeconomic status, exposure to multiple languages, developmental disorders)\
+* Yes / No\
+19. Do you have any concerns about your child’s language development?\
+* Yes / No
+
+**Part 6: Verify**
+
+20. Please select 'Agree' to verify that you are paying attention.\
+* Agree / Disagree\
+21. Do you agree to participate in this survey? Your responses will be kept confidential and used only for research purposes.\
+* Yes / No
+
+**End Part**
+
+Thank you for participating!
+
+Your responses are valuable to our research on early language development.
+If you have any questions or would like more information about our research, please feel free to contact us at [winniekeai23@gmail.com](mailto:winniekeai23@gmail.com)(Ziyuan Shen).
 

### 2024-12-04 paper/paper.qmd
update final paper

@@
 
-# CDI methodology overview and assessment
+# CDI methodology overview and improvement
 
@@
 
-## Survey implementation and Structure
+## Survey Implementation and Structure
 To reach a diverse audience of caregivers, multiple distribution channels will be used.
@@
 
-1. Is your child between the ages of 8 and 36 months?\
-* Yes / No\
-2. Are you the primary caregiver of the child?\
-* Yes / No\
-3. Are you willing to provide information about your child’s language development?\
-* Yes / No
+1.Is your child between the ages of 8 and 36 months?
+
+*Yes / No
+
+2.Are you the primary caregiver of the child?
+
+*Yes / No
+
+3.Are you willing to provide information about your child’s language development?
+
+*Yes / No
 
@@
 
-4. How old is your child?\
-* 8-12 months / 13-24 months / 25-36 months\
-5. What is your child's gender?\
-* Boy / Girl / Prefer not to say\
-6. What is your child’s race or ethnicity? (Other, please specify)\
-* White / Black / African American / Hispanic or Latino / Asian / Other\
-7. What is your highest level of education?(Other, please specify)\
-* Primary / Secondary / Some College / Bachelor’s Degree / Graduate Degree / Other\
-8. What is your current employment status? (Other, please specify)\
-* Employed / Unemployed / Student / Homemaker/ Other\
-9. What is your household income level?\
-* Less than $30,000 / $30,000-$59,999 / $60,000-$99,999 / Greater than $100,000 / Prefer not to say\
-10. How many language(s) do you speak at home?\
-* Monolingual (One language) / Bilingual (Two languages) / Multilingual (Three or more languages)\
+4.How old is your child?
+
+*8-12 months / 13-24 months / 25-36 months
+
+5.What is your child's gender?
+
+*Boy / Girl / Prefer not to say
+
+6.What is your child’s race or ethnicity? (Other, please specify)
+
+*White / Black / African American / Hispanic or Latino / Asian / Other
+
+7.What is your highest level of education?(Other, please specify)
+
+*Primary / Secondary / Some College / Bachelor’s Degree / Graduate Degree / Other
+
+8.What is your current employment status? (Other, please specify)
+
+*Employed / Unemployed / Student / Homemaker/ Other
+
+9.What is your household income level?
+
+*Less than $30,000 / $30,000-$59,999 / $60,000-$99,999 / Greater than $100,000 / Prefer not to say
+
+10.How many language(s) do you speak at home?
+
+*Monolingual (One language) / Bilingual (Two languages) / Multilingual (Three or more languages)
 
@@
 
-11. At what age did your child start saying their first word?\
-* 8-12 months / 13-24 months / 25-36 months\
-12. How often does your child use words or phrases (like “mama”, “more”)?\
-* Frequently (several times a day) / Occasionally (once or twice a day) / Rarely / Not yet\
-13. Does your child use gestures (like waving, pointing) to communicate?\
-* Yes, often / Occasionally / No\
-14. How would you rate your child’s vocabulary comprehension?\
-* Understands basic words (like names of family members) / Understands common household objects (like cup, ball) / Understands more complex phrases (like “Give me the red ball”) / Does not yet understand any words
+11.At what age did your child start saying their first word?
+
+*8-12 months / 13-24 months / 25-36 months
+
+12.How often does your child use words or phrases (such as “mama”, “more”)?
+
+*Frequently (several times a day) / Occasionally (once or twice a day) / Rarely / Not yet
+
+13.Does your child use gestures (such as waving, pointing) to communicate?
+
+*Yes, often / Occasionally / No
+
+14.How would you rate your child’s vocabulary comprehension?
+
+*Understands basic words (such as names of family members) / Understands common household objects (such as cup, ball) / Understands more complex phrases (such as “Give me the red ball”) / Does not yet understand any words
 
@@
 
-15. How often do you engage in activities that promote language development (like reading, or singing songs)?\
-* Daily / Several times a week / Once a week or less / Never\
-16. Do you use any of the following methods to support your child’s language development? (Check all that apply)\
-* Reading books aloud / Singing songs / Playing interactive games (like peek-a-boo) / Speaking to the child in multiple languages / None of the above\
-17. How often does your child interact with other children?\
-* Frequently / Occasionally / Rarely / Does not interact with other children
+15.How often do you engage in activities that promote language development (such as reading, or singing songs)?
+
+*Daily / Several times a week / Once a week or less / Never
+
+16.Do you use any of the following methods to support your child’s language development? (Check all that apply)
+
+*Reading books aloud / Singing songs / Playing interactive games (such as peek-a-boo) / Speaking to the child in multiple languages / None of the above
+
+17.How often does your child interact with other children?
+
+*Frequently / Occasionally / Rarely / Does not interact with other children
 
@@
 
-18. Do you believe there are any factors that have particularly influenced your child’s language development? (Such as: socioeconomic status, exposure to multiple languages, developmental disorders)\
-* Yes / No\
-19. Do you have any concerns about your child’s language development?\
-* Yes / No
+18.Do you believe there are any factors that have particularly influenced your child’s language development? (Such as: socioeconomic status, exposure to multiple languages, developmental disorders)
+
+*Yes / No
+
+19.Do you have any concerns about your child’s language development?
+
+*Yes / No
 
@@
 
-20. Please select 'Agree' to verify that you are paying attention.\
-* Agree / Disagree\
-21. Do you agree to participate in this survey? Your responses will be kept confidential and used only for research purposes.\
-* Yes / No
+20.Please select 'Agree' to verify that you are paying attention.
+
+*Agree / Disagree
+
+21.Do you agree to participate in this survey? Your responses will be kept confidential and used only for research purposes.
+
+*Yes / No
 

### 2024-12-04 scripts/05-exploratory_data_analysis.R
read_parquet

@@ -5,7 +5,7 @@
 # Contact: ziyuan.shen@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-#   - The `tidyverse`, `ggplot2` and `ggridges` packages must be installed and loaded
+#   - The `tidyverse`, `ggplot2`, `ggridges` and `arrow` packages must be installed and loaded
 #   - 03-clean_data.R must have been run
 # Any other information needed? Make sure you are in the `Factors_Influencing_Early_Childhood_Comprehension_of_American_English` rproj
 
@@ -14,9 +14,10 @@
 library(tidyverse)
 library(ggplot2)
 library(ggridges)
+library(arrow)
 
 #### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
 ## Set theme for better aesthetics
 theme_set(theme_minimal())
@@ -44,7 +45,7 @@ ggplot(analysis_data, aes(x = comprehension)) +
 ggplot(analysis_data, aes(x = production)) +
   geom_histogram(bins = 20, fill = "skyblue", color = "black") +
   labs(
-    title = "Distribution of Comprehension",
+    title = "Distribution of production",
     x = "production",
     y = "Count"
   )

### 2024-12-04 paper/paper.qmd
update final paper

@@
 ```{r}
+#| echo: false
 #| include: false
@@
 library(dplyr)
-library(here)
 library(ggplot2)
@@
 library(broom.mixed)
+
+analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
 ```
@@
 
+## Data Source
+The dataset used for this analysis is from the World Database of Children’s Vocabulary Growth [@citedata], specifically focusing on American English children.
+The dataset used in this study provides different data on the language comprehension skills of young children among American English language learners.
+The original dataset consists of 14,826 rows and 22 columns covering a range of early childhood developmental variables such as comprehension, output and caregiver education.
+
 ## Data Overview
@@
 We used the statistical programming language R [@citeR] to retrieve, simulate, clean, analyze, and test early childhood language comprehension data.
-The dataset used for this analysis is from the Wordbank database of children’s vocabulary growth [@citedata], specifically focusing on American English children.
-It contains comprehension and a variety of predictors, which include developmental, demographic, and environmental factors.
-Following the methodology discussed in “Telling Stories with Data” [@tellingstories], we
-discover how different predictors affect language comprehension in early childhood.
+The dataset contains comprehension and a variety of predictors, which include developmental, demographic, and environmental factors.
+Following the methodology discussed in “Telling Stories with Data” [@tellingstories], we discover how different predictors affect language comprehension in early childhood.
 
-The dataset used in this study provides different data on the language comprehension skills of young children among American English language learners.
-The dataset consists of 14,826 rows and 22 columns covering a range of early childhood developmental attributes such as comprehension, output and parental education.
-To ensure the validity of the analysis, we filtered the data to include children whose comprehension assessments and caregiver information were complete, which would ensure a general, unbiased representation of the target population.
+To ensure the validity of the analysis, we filtered the original dataset to include children whose comprehension assessments and caregiver information were complete, which would ensure a general, unbiased representation of the target population.
 This cleaning allowed us to focus on high-quality, well-integrated cases, which provided reliable results in young children's comprehension skills.
@@
 
-## Outcome variables
+## Outcome Variables
 
@@
 #| fig.height: 4
-
-analysis_data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
@@
 
-## Predictor variables
+## Predictor Variables
 There are eight predictor variables for my study after cleaning. The Descriptions of Each Predictor Variable is in @tbl-Predictor shows at [Appendix -@sec-data-descriptions].
@@
 
-## Model set-up
+## Model Set-up
 
@@
 
-## Model results and interpretation
+## Model Results and Interpretation
 
@@
 
-## Effect of birth order on language comprehension
+## Effect of Birth Order on Language Comprehension
 
@@
 
-# CDI methodology overview and improvement
+# CDI Methodology Overview and Improvement
 
@@
 
-## Target population, frame, and sample
+## Target population, Frame, and Sample
 
@@
 
-## Sample recruitment
+## Sample Recruitment
 
@@
 
-## Sampling approach and trade-offs
+## Sampling Approach and Trade-offs
 
@@
 
-## Non-response handling
+## Non-response Handling
 
@@
 
-### Survey structure
+### Survey Structure
 
-**Survey introduction:**
+**Survey Introduction:**
 
@@
 
-**Survey question:**
+**Survey Question:**
 
@@
 
-## Model summary {#sec-model-summary}
+## Model Summary {#sec-model-summary}
 
@@
 
-### Posterior predictive check
+### Posterior Predictive Check
 
@@
 \newpage
-### Gelman-Rubin diagnosis
+### Gelman-Rubin Diagnosis
 

### 2024-12-04 scripts/03-clean_data.R
update remove csv

@@ -58,7 +58,6 @@ analysis_data_train <- training(split)
 analysis_data_test <- testing(split)
 
 # Save cleaned data in different formats
-write_csv(cleaned_data, "data/02-analysis_data/analysis_data.csv")
 write_parquet(cleaned_data, "data/02-analysis_data/analysis_data.parquet")
 write_parquet(analysis_data_train, "data/02-analysis_data/train_data.parquet")
 write_parquet(analysis_data_test, "data/02-analysis_data/test_data.parquet")
